# Data sources

+ Standardized table of mhq refpoints, assessments and measurements

These tables contain the originally selected sampling units and the assessments and measurements.
We start from version 'rapportage2025'.

```{r}

path_mhq_terr <- "../../data/mhq_terr/rapportage2025"

mhq_terr_assessments <- read_vc("mhq_terr_assessments", 
                                root = path_mhq_terr)

mhq_terr_refpoints <- read_vc("mhq_terr_refpoints", 
                                root = path_mhq_terr) 
    
mhq_terr_measurements <- read_vc("mhq_terr_measurements", 
                                root = path_mhq_terr)

mhq_terr_popunits <- read_vc("mhq_terr_popunits", 
                                root = path_mhq_terr)
```

+ Administration INBO fieldwork coordinator

The INBO fieldwork coordinator uses [this googlesheet](https://docs.google.com/spreadsheets/d/1WkGPPYTb1Xytg5kJeOSxLvt3PUdFjlpPb4rwYG9znys/) to document which sampling units have been assessed.

In order to reproduce the workflow, we save a copy from the google sheet, dating 2026-08-24.
The file is saved in [this](https://drive.google.com/drive/folders/1wFZINPaR137csS3q79l8hpGkX24eEAaG) google drive folder.
Copy the file to the folder 'data/admin_mhq_terr' in this R project.

```{r}

assessments_admin_inbo_orig <- read_csv2("../../data/admin_mhq_terr/voortgang_mhq_terrestrisch_2026.csv")

```

TO DO Explain db_ref
TO DO user_reference instead of db_ref_orig?

```{r}

assessments_admin_inbo <- assessments_admin_inbo_orig %>%
    select(-starts_with("...")) %>%
    filter(!is.na(assessment_date)) %>%
    mutate(db_ref_orig = db_ref,
           fieldwork_team = str_to_lower(fieldwork_team),
           assessment_date = as.Date(assessment_date, format = "%d-%m-%Y"),
           grts_ranking_draw = as.numeric(grts_ranking_draw),
        db_ref = ifelse(nchar(db_ref_orig) > 5 & str_detect(db_ref_orig, "_"), str_sub(db_ref_orig, end = str_locate(db_ref_orig, "_") - 1), db_ref_orig)
       ) %>%
    filter(fieldwork_team == "inbo" | is.na(fieldwork_team)) %>%
    mutate(point_code = ifelse(point_code == "529473_2", "529473_1", point_code),
           type_observed = ifelse(point_code == "470546_1", "6230_ha", type_observed)) # CHANGE in admin table!

check_na <- assessments_admin_inbo %>%
    filter(is.na(lsvi_measurement))
#ok --> not measured yet

assessments_admin_inbo <- assessments_admin_inbo %>%
    filter(!is.na(lsvi_measurement))

```

+ assessments and measurement ANB

The assessment data and measurement data of ANB is exported from a Fieldmap database.

```{r}

path <- file.path(fileman_up("n2khab-mhq-data"), "processed/fieldmap_mhq")

status_admin_anb <- read_vc(root = path, file = "sample_status") 
assessments_admin_anb <- read_vc(root = path, file = "assessments") 

type_observed <- read_vc(root = path, file = "type_observed") 

coordinates_anb <- read_vc(root = path, file = "coordinates")  

```

+ Measurements INBOVEG

INBOVEG contains the actual collected data including the coordinates.

```{r}

path <- file.path(fileman_up("n2khab-mhq-data"), "processed/inboveg_mhq_terr")

header_mhq_terr <- read_vc(root = path, file = "header_mhq_terr") %>%
    mutate(assessment_date = as.Date(vague_date_begin))

classif_mhq_terr <- read_vc(root = path, file = "classif_mhq_terr") %>%
    select(-survey)

visits_mhq_terr <- header_mhq_terr %>%
    select(survey, user_reference, recording_givid, area, assessment_date, y = latitude, x = longitude) %>%
    left_join(classif_mhq_terr, by = "recording_givid") %>%
    select(-classif_key) %>%
    mutate(user_reference = ifelse(user_reference == "1474526_2", "1474526", user_reference), #Change in INBOVEG
           user_reference = ifelse(user_reference == "1346134_2", "1346134", user_reference), #Change in INBOVEG
           type_observed = ifelse(user_reference == "2541138" & is.na(type_observed), "7140_meso", type_observed),
           type_cover = ifelse(user_reference == "2541138" & is.na(type_cover), 100, type_cover)) %>% #Change in INBOVEG
    mutate(db_ref = str_to_lower(user_reference),
           db_ref = str_remove(db_ref, "double_"),
           db_ref = ifelse(str_detect(str_sub(db_ref, 1, 1), "[a-z]"), str_sub(db_ref, 1, 5), db_ref),
           db_ref = ifelse(nchar(db_ref) > 5 & str_detect(db_ref, "_"), str_sub(db_ref, end = str_locate(db_ref, "_") - 1), db_ref),
           db_ref = ifelse(str_sub(user_reference, 1, 2) == "BV", str_to_lower(user_reference), db_ref),
           year = year(assessment_date)
           )

check <- visits_mhq_terr %>%
    group_by(user_reference, area, assessment_date, type_observed, y , x) %>%
    filter(n() > 1)

visits_mhq_terr <- visits_mhq_terr %>%
    filter(!str_detect(user_reference, "double"))

check_missing_area <- visits_mhq_terr %>%
    filter(is.na(area))

check_missing_coord <- visits_mhq_terr %>%
    filter(x == 500 | y == 500)

visits_mhq_terr_wide <- visits_mhq_terr %>%
    mutate(cover_description = str_c(ifelse(is.na(type_cover), "?", type_cover), "% ", type_observed)) %>%
    mutate(plot_type = ifelse(area <= 9 | is.na(area), "square", "circle")) %>%
    mutate(plot_type = ifelse(recording_givid == "IV2025071010094321", "circle", plot_type)) %>% #!CHANGE IN INBOVEG
    group_by(survey, db_ref, user_reference, recording_givid, plot_type, assessment_date, year, x, y) %>%
    summarise(cover_description = str_c(cover_description, collapse = "; ")) %>%
    ungroup() %>%
    st_as_sf(coords = c("x", "y"), crs = 4326) %>%
    st_transform(31370)

check <- visits_mhq_terr_wide %>%
    group_by(db_ref, user_reference, assessment_date, plot_type) %>%
    filter(n() > 1)

check2 <- visits_mhq_terr_wide %>%
    st_drop_geometry() %>%
    group_by(db_ref, survey, assessment_date) %>%
    mutate(n = n_distinct(user_reference)) %>%
    filter(n > 1)

visits_mhq_terr_wide <- visits_mhq_terr_wide %>%
    mutate(x_measured = round(st_coordinates(visits_mhq_terr_wide)[,1], 2),
           y_measured = round(st_coordinates(visits_mhq_terr_wide)[,2], 2)) %>%
    st_drop_geometry() %>%
    mutate(recording_givid = str_c(plot_type, ": ", recording_givid)) %>%
    group_by(db_ref, user_reference, assessment_date, year) %>%
    mutate(recording_givid = str_c(recording_givid, collapse = "; "),
           x_measured = mean(x_measured),
           y_measured = mean(y_measured)) %>%
    ungroup() %>%
    pivot_wider(names_from = "plot_type", values_from = "cover_description") 

mhq_terr_measurements <- mhq_terr_measurements %>%
    mutate(assessment_date = measurement_date)

visits_mhq_terr_new <- visits_mhq_terr_wide %>%
    anti_join(mhq_terr_measurements, by = c("user_reference", "assessment_date")) %>%
    anti_join(mhq_terr_measurements, by = c("db_ref", "assessment_date")) %>%
    filter(!is.na(square))

check_unique <- visits_mhq_terr_new %>%
    group_by(assessment_date, db_ref) %>%
    filter(n() > 1)

check_na_square <- visits_mhq_terr_wide %>%
    filter(is.na(square)) %>%
    filter(str_detect(survey, "N2000"))

# 36790 square measurement zit in survey 'N2000_grasland_fieldapp_3' --> nog toevoegen aan measurements

```

# Update assessments

## Assessments INBO

```{r}
assessments_table_old <- mhq_terr_assessments %>%
    mutate(grts_ranking_draw = as.numeric(str_sub(point_code, end = -3))) %>%
    left_join(mhq_terr_refpoints, by = "point_code") %>%
    mutate(year = year(assessment_date))

assessments_admin_inbo_new <- assessments_admin_inbo %>%
    mutate(year = year(assessment_date)) %>%
    anti_join(assessments_table_old, by = c("point_code", "assessment_date")) %>%
    left_join(select(assessments_table_old, point_code, year, x_exist = x, y_exist = y, assessment_date_exist = assessment_date), by = c("year", "point_code")) %>%
    mutate(dist = ((x_exist - x)^2 + (y_exist - y)^2)^0.5)

# here we check (unexpected) new assessments before 2025
check_new <- assessments_admin_inbo_new %>%
    filter(year < 2025)

```

Interpretation:

+ 1085366:
    + 1085366_1: selected for 2330_bu (ANB) and for 6230_hn (INBO); visited by ANB on 20-7-2017 and measured; visited by INBO on 31-7-2017 and not measured but replaced to 1085366_2; INBO assessment of 1085366_1 can be added

+ 1381569: similar situation as for 1085366

+ 50102: 50102_1 was visited twice in cycle 2, in 2019 and in 2022. 2019 assessment can be added.

+ 529473_1: to be added, probably missing due to wrong point_code in admin table (529473_2) --> corrected in admin table

+ 379222_1: to be added, probably missing due to wrong point_code (379222) --> corrected in admin table

Conclusion: all detected new assessments before 2025 can be added to the updated tables 

### Link new assessments to measurements INBO

+ new assessments with a matching new measurement in INBOVEG based on assessment date and user_reference (db_ref)

```{r}

assessments_measured <- assessments_admin_inbo_new %>%
    filter(lsvi_measurement) %>%
    inner_join(visits_mhq_terr_new, by = c("db_ref", "year"), suffix = c("_admin", "_inboveg"))

check_date <- assessments_measured %>%
    filter(assessment_date_admin != assessment_date_inboveg)

assessments_measured <- assessments_admin_inbo_new %>%
    filter(lsvi_measurement) %>%
    inner_join(visits_mhq_terr_new, by = c("db_ref", "assessment_date", "year"))
```

+ new assessments but no matching measurement in INBOVEG based on assessment date and user_reference (db_ref) --> OK

```{r}
assessments_measured_notinboveg <- assessments_admin_inbo_new %>%
    filter(lsvi_measurement) %>%
    anti_join(visits_mhq_terr_wide, by = c("db_ref", "assessment_date")) 

```

```{r}
assessments_measured_notinboveg %>%
    select(point_code, db_ref, assessment_date, lsvi_measurement, type_observed) %>%
    datatable(rownames = FALSE, filter = "top")
```

+ new assessments that are not measured (and therefore no record in INBOVEG) --> OK

```{r}
assessments_not_measured <- assessments_admin_inbo_new %>%
    filter(!lsvi_measurement)
    
#check if there is no inboveg record

check <- assessments_not_measured %>%
    semi_join(visits_mhq_terr_wide, by = c("db_ref", "assessment_date"))
# ok: gaat om de originele locaties van verplaatste punten

check_old <- assessments_not_measured %>%
    filter(year < 2025)
# ok
```

+ new inboveg records with no assessments records in the admin table --> TO DO: add 1474526_2 in admin table

```{r}

assessments_add <- visits_mhq_terr_new %>%
    anti_join(assessments_admin_inbo_new, by = c("db_ref", "assessment_date")) %>%
    anti_join(assessments_admin_inbo, by = c("db_ref", "assessment_date")) %>%
    #filter(!str_detect(db_ref, "bv")) %>%
    filter(!str_detect(db_ref, "double|000000006")) %>%
    filter(!str_detect(db_ref, "dummy")) %>%
    filter(str_detect(survey, "N2000")) %>%
    filter(!str_detect(db_ref, "iv"))

```

```{r}
assessments_add %>%
    datatable(filter = "top",
              rownames = FALSE)
```
Interpretation: 

+   923346: assessment in 2025 is incorrect: ignore
 
+   1474526!!: 
    +  add record in admin table for 2025 for 1474526_2

+ 2023 records for 627922 and 857298 are not correct: ignore. 


+ new inboveg records for existing records in the admin table --> TO DO add measurements in mhq_terr_measurements

```{r}

assessments_table_old <- assessments_table_old %>%
    mutate(user_reference = str_sub(point_code, end = -3))

measurements_add <- visits_mhq_terr_new %>%
    anti_join(assessments_admin_inbo_new, by = c("db_ref", "assessment_date")) %>%
    semi_join(assessments_table_old, by = c("user_reference", "assessment_date")) %>%
    filter(!str_detect(db_ref, "bv")) %>%
    filter(!str_detect(db_ref, "double")) %>%
    filter(str_detect(survey, "N2000")) %>%
    filter(!str_detect(db_ref, "iv"))
#check
```


```{r}
measurements_add %>%
    datatable(rownames = FALSE,
              filter = "top")
```


Interpretation:

+ 6986542_2, 6630337_2, 7678913_2: in lsvi_measurement = FALSE in admin table --> corrected 

+ 6986542_2, 6630337_2, 7678913_2, 6572681_2, 207406_2:
    + measurement missing in mhq_terr_measurements
    + all replacements where another type is measured than the target type (invalid measurements) 
    + none of the sampling units are included in cycle 2

+ 4008897, 5295334, 606390:
    + measurement missing in mhq_terr_measurements
    + measurements in 2024

+ others

```{r}
assessments_admin_other <- assessments_admin_inbo_new %>%
    anti_join(assessments_measured, by = c("db_ref", "assessment_date")) %>%
    anti_join(assessments_not_measured, by = c("db_ref", "assessment_date")) %>%
    anti_join(assessments_measured_notinboveg, by = c("db_ref", "assessment_date")) %>%
    anti_join(assessments_add, by = c("db_ref", "assessment_date"))

```

+ Let's check if the observed type in the admin table is missing. 

```{r}
check_type_observed <- assessments_measured %>%
    filter(is.na(type_observed)) %>%
    mutate(one_type = str_sub(square, 1, 3) == "100")

check_one_type <- all(check_type_observed$one_type)


```

```{r}

if (nrow(check_type_observed) > 0) {
    
  check_type_observed %>%
    select(point_code, db_ref, assessment_date, lsvi_measurement, type_observed, square, circle) %>%
    kable() %>%
    kable_styling()  
} else {
    
    cat("OK!")
}

```

```{r}
assessments_admin_inboveg <- assessments_measured %>%
    bind_rows(assessments_not_measured) %>%
    select(fieldwork_team, point_code, db_ref, reassessment_1, relocation, grts_relocation, assessment_date, inaccessible, lsvi_measurement, type_observed,
           type_target_all, x_measured, y_measured, year, survey, recording_givid, user_reference, circle, square) %>%
    mutate(point_code = str_trim(point_code))

# To be changed in admin table

assessments_admin_inboveg <- assessments_admin_inboveg %>%
    mutate(point_code = ifelse(!str_detect(point_code, "_"), 
                               str_c(point_code, "_1"),
                                     point_code))
```

### Link assessments to refpoints

Next we link the new assessments to the refpoints.

```{r}

refpoints_assessed_inbo_new <- mhq_terr_refpoints %>%
    inner_join(assessments_admin_inboveg, by = "point_code") %>%
    distinct() %>%
    mutate(distance = round(sqrt((x_measured - x)^2 + (y_measured - y)^2), 3))

check_missing_refpoint <- assessments_admin_inboveg %>%
    anti_join(refpoints_assessed_inbo_new, by = "point_code")

check_double_su <- refpoints_assessed_inbo_new %>%
    group_by(point_code, type_target_all, assessment_date) %>%
    mutate(n_pointcode = n()) %>%
    ungroup() %>%
    filter(n_pointcode > 1)

```

+ check if refpoints occur in sampling plan cycle 2

```{r}
# ADD this code to design repo

grts_master <- read_GRTSmh()

su_duinen_cyclus2 <- read.csv2("../../data/design_cycle2/mhq_terr_cyclus2_duinen_inbo_2023-05-11.csv")

su_anb_cyclus2 <- read.csv2("../../data/design_cycle2/mhq_terr_cyclus2_anb.csv")

su_inbo_cyclus2 <- read.csv2("../../data/design_cycle2/mhq_terr_cyclus2_inbo.csv")

su_forest <- read.csv2("../../data/design_cycle1/mhq_forests_cycle1.csv") %>%
    mutate(grts_ranking_draw = grts_ranking)

sampling_plan_inbo_c2 <- su_duinen_cyclus2 %>%
    bind_rows(su_inbo_cyclus2) %>%
    separate_wider_delim(point_code, delim = "_", names = c("point_code_base", "point_code_suffix"), cols_remove = FALSE)

sampling_plan_c2 <- su_duinen_cyclus2 %>%
    bind_rows(su_inbo_cyclus2) %>%
    bind_rows(su_anb_cyclus2) %>%
    mutate(reassessment = type_meetpunt == "bestaand meetpunt",
           oversampling_subtype = uitvoer_opname == "subtype aanwezig") %>%
    select(panel, point_code, polygon_id, grts_ranking_draw, sac = sbz, main_type = hoofdtype, type = habitattype, phab, reassessment, date_assessment_c1 = datum_opname_cyclus1, oversampling_subtype, db_ref,  recording_givid, x, y)

su_duinen_vervangpunten_cyclus2 <- read.csv2("../../data/design_cycle2/mhq_terr_vervangpunten_duinen_inbo_2023-05-11.csv")

su_anb_cyclus2 <- read.csv2("../../data/design_cycle2/mhq_terr_vervangpunten_anb.csv")

su_inbo_cyclus2 <- read.csv2("../../data/design_cycle2/mhq_terr_vervangpunten_inbo.csv")

relocation_points <- su_duinen_vervangpunten_cyclus2 %>%
    bind_rows(su_anb_cyclus2) %>%
    bind_rows(su_inbo_cyclus2) %>%
    select(point_code_relocation = point_code_vervang, polygon_id, rank_rel_buffer, rank_rel_polygon, x, y) %>%
    st_as_sf(coords = c("x", "y"), crs = 31370, remove = FALSE)

relocation_points <- relocation_points %>%
    mutate(grts_ranking = (terra::extract(grts_master, relocation_points))$GRTSmaster_habitats) %>%
    st_drop_geometry() %>%
    select(point_code_relocation, polygon_id, grts_ranking, rank_rel_buffer, rank_rel_polygon, x, y) 

```

```{r}
check_su2 <- refpoints_assessed_inbo_new %>%
    separate_wider_delim(point_code, delim = "_", names = c("point_code_base", "point_code_suffix"), cols_remove = FALSE) %>%
    filter(!point_code %in% sampling_plan_inbo_c2$point_code) %>%
    mutate(check_point_code_base = point_code_base %in% sampling_plan_inbo_c2$point_code_base) %>%
    left_join(sampling_plan_inbo_c2 %>%
                  select(point_code_base, point_code_sp = point_code), by = "point_code_base")

# ok: only assessments from cycle 1
```

+ We check if metadata is missing

```{r}
check_metadata <- refpoints_assessed_inbo_new %>%
    filter(lsvi_measurement) %>%
    filter(is.na(reassessment_1) | is.na(relocation))

# To be changed in admin table!
refpoints_assessed_inbo_new <- refpoints_assessed_inbo_new %>%
    mutate(relocation = ifelse(point_code == "326965_2", FALSE, relocation))
```


+ Check if there are assessments with relocation = FALSE but where the measured location is > 1 meter away from the reference point

```{r}

check_coordinates <- refpoints_assessed_inbo_new %>%
    filter(distance >= 1, !relocation)

if (nrow(check_coordinates) > 0) {
    
    check_coordinates %>%
    select(point_code, db_ref, assessment_date, type_observed, x, y, x_measured, y_measured, distance, relocation, reassessment_1) %>%
    datatable(rownames = FALSE, filter = "top")

check_coordinates %>%
    write_csv2("../../output/check_coördinaten.csv")
    
} else {
    
    cat("OK!")
    
}

```

## Assessments ANB

### Assessments in Fieldmap

```{r}

#waarom deze stap?
mhq_terr_assessments <- mhq_terr_assessments %>%
    mutate(db_ref = str_sub(point_code, end = -3))

coordinates_anb <- coordinates_anb %>%
    mutate(plot_id = as.character(plot_id))  

assessments_admin_anb_new <- assessments_admin_anb %>%
    mutate(db_ref = plot_id) %>%
    filter(type_target_all != "91E0_sf",
           is.na(remark) | remark != "opname door INBO ikv verdroging") %>%
    rename(assessment_date = date_assessment) %>%
    anti_join(mhq_terr_assessments, by = c("db_ref", "assessment_date")) 
```

+ assessments in the second monitoring cycle: point_code

```{r}
coordinates_anb_c2 <- coordinates_anb %>%
    filter(mon_cycle == 2) %>%
    select(point_code, x_measured, y_measured, mon_cycle, relocation = verlegd, grts_relocation = nr_vervangplot)

check_unique <- coordinates_anb_c2 %>%
    group_by(point_code) %>%
    filter(n() > 1)
# OK

assessments_anb_c2 <- assessments_admin_anb_new %>%
    filter(mon_cycle == 2) %>%
    rename(point_code = sampling_unit_code) %>%
    filter(remark != "opname door INBO ikv verdroging" | is.na(remark)) %>%
    left_join(coordinates_anb_c2, by = c("point_code", "mon_cycle"))
    
refpoints_assessed_anb_c2 <- mhq_terr_refpoints %>%
    inner_join(assessments_anb_c2, by = "point_code") %>%
    mutate(distance = round(sqrt((x_measured - x)^2 + (y_measured - y)^2), 3)) %>%
    mutate(relocation = ifelse(is.na(relocation) & distance < 1, FALSE, relocation))

check_missing_refpoint <- assessments_anb_c2 %>%
    anti_join(refpoints_assessed_anb_c2, by = "point_code")
#OK
```

## Combine

We combine the new assessments of ANB and INBO and do some quality checks.

```{r}

refpoints_assessed_all <- refpoints_assessed_inbo_new %>%
    mutate(not_measurable = NA,
           assessment_source = "field assessment") %>%
    bind_rows(refpoints_assessed_anb_c2 %>%
                   mutate(fieldwork_team = "anb",
                          survey = "anb_c2")) %>%
    select(survey, fieldwork_team, point_code, is_centroid, grts_ranking, type_target_all, type_observed, x, y, db_ref, assessment_source, assessment_date, inaccessible, not_measurable, lsvi_measurement, relocation, grts_relocation, x_measured, y_measured, user_reference, recording_givid, distance, remark) %>%
    arrange(grts_ranking) %>%
    mutate(type_target_all = ifelse(type_target_all == "2190", "2190_overig", type_target_all))
```

+ Is the subtype of the observed habitat type defined? If the observed subtype is not defined, we assume that the subtype corresponds with the target subtype.

```{r}

types <- read_types()

has_subtypes <- types %>%
    group_by(main_type) %>%
    summarise(has_subtype = n() > 1 ) %>%
    ungroup() %>%
    filter(has_subtype) %>%
    filter(main_type != "9120") %>%
    filter(!str_detect(main_type, "rbb"))

check_subtype <-  refpoints_assessed_all %>%
    filter(type_observed %in% has_subtypes$main_type)

if (nrow(check_subtype) > 0) {
    
    add_subtype_observed <- refpoints_assessed_all %>%
    filter(type_observed %in% has_subtypes$main_type) %>%
    separate(type_target_all, into = c("type_target_1", "type_target_2"), sep = ";", remove = FALSE) %>%
    mutate(type_observed = ifelse(str_sub(type_target_1, 1, 4) == type_observed, type_target_1,
                                  ifelse(str_sub(type_target_2, 1, 4) == type_observed & !is.na(type_target_2), type_target_2, type_observed))) %>%
    select(-type_target_1, -type_target_2)

refpoints_assessed_all <- refpoints_assessed_all %>%
    anti_join(add_subtype_observed, by = c("point_code", "assessment_date")) %>%
    bind_rows(add_subtype_observed)

} else {
    
    cat("OK: all subtypes are defined")
}

```
+ Check if the status of the assessment is defined, this is the case when we either have a value for `assessment_source`, `inaccessible` or `not_measurable`

```{r}
# status of each survey event defined?
check_status <- refpoints_assessed_all %>%
    filter(is.na(assessment_source) & is.na(inaccessible) & is.na(not_measurable))

if (nrow(check_status) > 1) {
    
    cat("CHeck following records")
    
    check_status %>%
        datatable(rownames = FALSE)
    
} else {
    
    cat("OK: status is defined for all records")
    
}

```

+ Check if sampling units are assessed by both ANB and INBO. If this is the case we select the assessment with a measurement.

```{r}
check_unique_point_code <- refpoints_assessed_all %>%
    mutate(year = year(assessment_date)) %>%
    group_by(point_code, year) %>%
    filter(n() > 1)

if (nrow(check_unique_point_code) > 1) {
    
    check_unique_point_code %>%
        datatable(rownames = FALSE)
    
    refpoints_assessed_all <- refpoints_assessed_all %>%
    mutate(year = year(assessment_date)) %>%
    group_by(point_code, year) %>%
    filter(n() == 1 | (n() > 1 & lsvi_measurement)) %>%
    ungroup()
    
} else {
    
    cat("OK: no sampling units are assessed both by ANB and INBO")
    
}

```

# Create new reference points for relocated population units

When a sampling unit is relocated, we have to define a new reference point (`point_code`).
Later we will check if the relocation is performed according to the protocol.

First we check the distance between the sample unit reference point and the measured coordinates.
When indicated that the assessment was not relocated, this distance should be smaller than 1 m (we allow a 1 meter location error). 

```{r}
check_no_replacement <- refpoints_assessed_all %>%
    filter(!relocation) %>%
    filter(distance > 1)
```

```{r}
refpoints_assessed_all %>%
    filter(lsvi_measurement) %>%
    ggplot(aes(x = distance)) +
    geom_histogram() +
    facet_wrap(~str_c("relocation = ",relocation), scales = "free", ) +
    labs(x = "Aantal meetpunten")

```

Next we determine the `grts_ranking` of the new location and the distance to the nearest grid cell centre.
This distance should be smaller than 1 m.

```{r}

distance_from_centroidgrid <- function(min_x, min_y, measured_x, measured_y, resolution = 32, threshold = 1) {
    
    difference_x <- (measured_x - min_x) %% resolution
    shortest_distance_x <- min(difference_x, resolution - difference_x)
    difference_y <- (measured_y - min_y) %% resolution
    shortest_distance_y <- min(difference_y, resolution - difference_y)
    distance_from_grid <- sqrt( shortest_distance_x ^ 2 + shortest_distance_y ^ 2)
    
    return(distance_from_grid)
}

min_x <- (mhq_terr_refpoints %>%
    filter(is_centroid) %>%
    filter(x == min(x)))$x

min_y <- (mhq_terr_refpoints %>%
    filter(is_centroid) %>%
    filter(y == min(y)))$y

new_reference_points <- refpoints_assessed_all %>%
    filter(distance >= 1) %>%
    filter(lsvi_measurement) %>%
    group_by(point_code) %>%
    mutate(ranking_code = str_sub(point_code, end = -3),
           x_orig = x,
           y_orig = y,
           x = x_measured,
           y = y_measured,
           distance_centroid = distance_from_centroidgrid(min_x = min_x,
                              min_y = min_y,
                              measured_x = x_measured,
                              measured_y = y_measured),
           is_centroid = distance_centroid < 1,
           legacy_site = FALSE) %>%
    ungroup() %>%
    st_as_sf(coords = c("x", "y"), crs = 31370)

new_reference_points <- new_reference_points %>%
    mutate(grts_ranking = (terra::extract(grts_master,new_reference_points))$GRTSmaster_habitats) %>%
    st_drop_geometry() 
```

```{r}
check_relocation <- new_reference_points %>%
    filter(distance_centroid > 1)

if (nrow(check_relocation) > 0) {
    
    cat("Check relocations below")
    
    datatable(check_relocation,
              rownames = FALSE)
} else {
    
    cat("OK: all relocations are within a distance of 1 m from a grid cell centroid")
}
```

Finally we define a new point code for the relocated sampling unit as follows:

+ the point_code prefix (= `grts_ranking_draw`) of the original point_code is kept
+ a new point_code suffix is created by augmenting the maximum suffix value (for the point_code prefix) by one

We also check if the new point_code does not already exist.

```{r}

max_n <- mhq_terr_refpoints %>%
    select(point_code) %>%
    separate(col = point_code, into = c("left_part", "right_part"), sep = "_", remove = FALSE) %>%
    mutate(right_part = as.numeric(right_part)) %>%
    group_by(left_part) %>%
    mutate(max_n_point_code = max(right_part)) %>%
    ungroup() %>%
    select(point_code, max_n_point_code) %>%
    unique()

new_reference_points <- new_reference_points %>%
    left_join(max_n, by = "point_code") %>%
    mutate(x = x_measured,
           y = y_measured) %>%
    group_by(ranking_code) %>%
    mutate(n = n(),
           rank = rank(x),
           add = rank(x) + max_n_point_code,
           point_code_old = point_code,
          point_code = str_c(ranking_code, "_", add)
        ) %>%
    ungroup()

# check ik new reference points already exist
check <- new_reference_points %>%
    semi_join(mhq_terr_refpoints, by = "point_code")

original_reference_points <- refpoints_assessed_all %>%
    semi_join(new_reference_points, by = c("point_code" = "point_code_old", "assessment_date")) %>%
    mutate(lsvi_measurement = FALSE,
           type_observed = "gdh", #geen doelhabitat
           distance = 0,
           x_measured = NA,
           y_measured = NA,
           user_reference = NA,
           recording_givid = NA,
           relocation = FALSE) %>%
    left_join(select(new_reference_points, point_code = point_code_old, relocated_to = point_code),
              by = "point_code")

check <- original_reference_points %>%
    group_by(point_code) %>%
    filter(n() > 1)

refpoints_assessed_all <- refpoints_assessed_all %>%
    anti_join(original_reference_points, by = c("point_code", "assessment_date")) %>%
    bind_rows(original_reference_points) %>%
    bind_rows(new_reference_points)

check_unique_point_code <- refpoints_assessed_all %>%
    group_by(point_code, assessment_date) %>%
    filter(n() > 1)

check_replaced <- refpoints_assessed_all %>%
    filter(relocation) %>%
    filter(str_detect(point_code, "_1"))

```
