# Validation

## Visualise assessments

```{r}

new_assessments_sf <- refpoints_assessed_all %>%
    arrange(assessment_date) %>%
    mutate(grts_relocation = ifelse(is.na(grts_relocation), FALSE, grts_relocation)) %>%
    st_as_sf(coords = c("x", "y"), crs = 31370) %>%
    mutate(status = ifelse(!is.na(inaccessible), str_c("Inaccesible: ", inaccessible),
                           ifelse(!is.na(not_measurable), str_c("Not measurable: ", not_measurable),
                                  ifelse(!lsvi_measurement, "Not measured",
                                         ifelse(grts_relocation, "Measured with relocation", "Measured without relocation")))),
           status = factor(status)) %>%
    select(assessment_date, point_code, type_target_all, type_observed, status) %>%
    st_transform(crs = 4326)

pal <- colorFactor(
  palette = c("red", "orange", "darkgreen", "yellow", "black", "blue"),
  domain = new_assessments_sf$status
)

sd <- SharedData$new(new_assessments_sf)

filter_checkbox("status", "Status meetpunt", sd, ~status)

sd %>%
    leaflet() %>%
    addTiles() %>%
    addCircleMarkers(color = ~pal(status), popup = ~str_c(point_code, ": ", type_target_all, assessment_date)) %>%
    addLegend(pal = pal, values = ~status)

datatable(sd, 
          rownames = FALSE,
          filter = "top"
          )
```

## Observed habitat type

+ Does the observed habitat type correspond with the target habitat type?

```{r}

max_ranking_sample <- sampling_plan_c2 %>%
    group_by(type, sac) %>%
    summarise(max_ranking = max(grts_ranking_draw)) %>%
    ungroup() %>%
    rename(type_observed = type)

measured_check <- refpoints_assessed_all %>%
    filter(lsvi_measurement) %>%
    mutate(point_code_old = ifelse(is.na(point_code_old), point_code, point_code_old)) %>%
    select(point_code, point_code_old, assessment_date, type_target_all, type_observed, remark) %>%
    filter(type_observed != type_target_all) %>%
    left_join(sampling_plan_c2 %>%
                  select(point_code_old = point_code, main_type, reassessment, oversampling_subtype, sac, grts_ranking_draw),
              by = "point_code_old") %>%
    mutate(valid_measurement = ifelse(reassessment, TRUE,
                               ifelse(oversampling_subtype, type_observed == type_target_all,
                                      str_sub(type_observed, 1, 4) == main_type))) %>%
    left_join(max_ranking_sample, by = c("sac", "type_observed"))
```

The field protocol has following rules:

+ sampling unit is being reassessed (i.e. was measured in cycle 1): always do a measurement even if the observed type is different than the target type, unless the land cover has clearly changed, do not relocate sampling unit 
+ new sampling unit
    + sampling unit was selected for main type: observed main type should match with target main type, if not relocate sampling unit within habitat polygon
    + oversampling for subtype: observed subtype should match target subtype, if not relocate sampling unit within habitat polygon

Following measurements are not according to these rules:

```{r}
measured_check %>%
    filter(!valid_measurement) %>%
    select(-point_code, -main_type, -sac, -grts_ranking_draw, -max_ranking) %>%
    datatable(rownames = FALSE)
```

+ For 251358_1 we assume that the type_observed corresponds with the target type -> valid measurement

```{r}
refpoints_assessed_all <- refpoints_assessed_all %>%
    mutate(type_observed = ifelse(point_code == "251358_1", "4010", type_observed))
```

+ For 255454_1 the observed type (4030) is different than the target type (4010) and this also is the case for all available relocation points (see remark).
This might also be the case for 29302_1 and 166102_1.
Let's check if the sampling unit is valid based on `grts_ranking_draw`

```{r}
measured_check %>%
    filter(!valid_measurement) %>%
    select(point_code_old, type_target_all, type_observed, grts_ranking_draw, max_ranking) %>%
    mutate(valid_sampling_unit = grts_ranking_draw < max_ranking) %>%
    datatable(rownames = FALSE)
```

## Relocations

In case of a relocation we check if a valid relocation point is selected and if the relative ranking of the relocation point recorded by the field worker (rank_rel_field) is correct.

```{r}
relocation_check <- refpoints_assessed_all %>%
    filter(relocation, distance > 0) %>% # distance = 0 --> original not measured location
    select(point_code_old, assessment_date, lsvi_measurement, type_target_all, type_observed, grts_ranking, grts_relocation, distance, is_centroid, x_measured, y_measured) %>%
    left_join(relocation_points %>%
                  rename(point_code_old = point_code_relocation),
              by = c("point_code_old", "grts_ranking"))
```

In the table below we show all relocations.

```{r}
relocation_check %>%
    select(point_code_old, assessment_date, type_target_all, type_observed, distance, is_centroid, rank_rel_field = grts_relocation, rank_rel = rank_rel_buffer) %>%
    datatable(rownames = FALSE,
              filter = "top")
```
```{r}
relocation_not_valid <- relocation_check %>%
    filter(is.na(rank_rel_buffer))

if (nrow(relocation_not_valid) == 0) {
    
    cat("For all relocations valid relocation points are selected!")
} else {
    
    cat("Following relocations are not valid. \n")
    
    datatable(relocation_not_valid) 
    
}

```

```{r}
relocation_check_admin <- relocation_check %>%
    filter(rank_rel_buffer != grts_relocation)

if (nrow(relocation_check_admin) == 0) {
    
    cat("For all relocations the relative rank recorded in the field matches the relative rank of the measured sampling unit!")
} else {
    
    cat("For following relocations the relative rank recorded in the field does not match the relative rank of the measured sampling unit. \n")
    
    relocation_check_admin %>%
        select(point_code_old, assessment_date, type_observed, distance, rank_rel_field = grts_relocation, rank_rel = rank_rel_buffer) %>%
        datatable(rownames = FALSE,
              filter = "top")
    
}

```

```{r, eval = FALSE}

habmap_stdized <- read_habitatmap_stdized(file = file.path(fileman_up("n2khab_data"), "/20_processed/_versions/habitatmap_stdized/habitatmap_stdized_2020_v1/habitatmap_stdized.gpkg"),
                                          version = "habitatmap_stdized_2020_v1") 

refpoint_select <- "9170737_1"

su_select <- refpoints_assessed_all %>%
    filter(point_code == refpoint_select) %>%
    st_as_sf(coords = c("x", "y"), crs = 31370) %>%
    st_transform(4326)

polygon_id_select <- sampling_plan_c2 %>%
    filter(point_code == refpoint_select)

polygon_select <- habmap_stdized$habitatmap_polygons %>%
    filter(polygon_id == polygon_id_select$polygon_id) %>%
    st_transform(4326)

measured_select <- relocation_check %>%
    filter(point_code_old == refpoint_select) %>%
    st_as_sf(coords = c("x_measured", "y_measured"), crs = 31370) %>%
    st_transform(4326)

relocation_points_show <- relocation_points %>%
    filter(point_code_relocation == refpoint_select) %>%
    st_as_sf(coords = c("x", "y"), crs = 31370) %>%
    st_transform(4326)

su_select %>%
    leaflet() %>%
    addProviderTiles(providers$Esri.WorldImagery) %>%
    addPolygons(data = polygon_select) %>%
    addCircleMarkers(color = "orange") %>%
    addCircleMarkers(data = measured_select, color = "blue") %>%
    addCircleMarkers(data = relocation_points_show, color = "yellow", radius = 1, label = ~rank_rel_buffer,
                     labelOptions = labelOptions(noHide = TRUE))
```


## Reassessments

In case sampling units that have been measured in cycle 1 are being reassessed, we expect to observe the same type as observed in cycle 1.
If not, an lsvi-measurement is performed even tough the observed type does not match the target type.
The idea is to keep these sampling units in the sample frame as the change in habitat type might be caused by a temporal disturbance (for example drought), which causes that certain key species are absent.
However when land cover or land use clearly changed, the sampling unit will not bet kept.
This should be interpreted by a habitat expert. 

```{r}
reassessment_check <- refpoints_assessed_all %>%
    left_join(sampling_plan_c2 %>%
                  select(point_code, main_type, reassessment, oversampling_subtype, sac, grts_ranking_draw),
              by = "point_code") %>%
    mutate(type_match = type_target_all == type_observed,
           main_type_match = str_sub(type_target_all, 1, 4) == str_sub(type_observed, 1, 4),
           status = ifelse(lsvi_measurement, "lsvi measurement",
                           ifelse(!is.na(inaccessible), str_c("inaccessible: ", inaccessible),
                                  ifelse(!is.na(not_measurable), str_c("not_measurable: ", not_measurable), "no lsvi measurement")))) %>%
    filter(reassessment) %>%
    select(point_code, assessment_date, status, lsvi_measurement, type_target_all, type_observed, type_match, distance, remark)
```


```{r}
reassessment_check %>%
    ggplot(aes(x = type_match, fill = status)) +
    geom_bar()
    
```

Below an overview of all assessments where no measurement is performed.
We expect a measurement for the sampling units that are accessible and measurable.
To be checked by habitat expert.

```{r}
reassessment_check %>%
    filter(!lsvi_measurement) %>%
    datatable(rownames = FALSE,
              filter = "top")
    
```
Below an overview of all assessments where a measurement is performed even tough the observed type is different than the target type.
Do we keep these sampling units?.
To be checked by habitat expert.

```{r}
reassessment_check %>%
    filter(lsvi_measurement) %>%
    filter(!type_match) %>%
    select(point_code, assessment_date, status, type_target_all, type_observed, type_match) %>%
    datatable(rownames = FALSE,
              filter = "top")
    
```

## New sampling units

```{r}
new_su_check <- refpoints_assessed_all %>%
    left_join(sampling_plan_c2 %>%
                  select(point_code, main_type, reassessment, oversampling_subtype, sac, grts_ranking_draw),
              by = "point_code") %>%
    mutate(type_match = type_target_all == type_observed,
           main_type_match = str_sub(type_target_all, 1, 4) == str_sub(type_observed, 1, 4),
           status = ifelse(lsvi_measurement, "lsvi measurement",
                           ifelse(!is.na(inaccessible), str_c("inaccessible: ", inaccessible),
                                  ifelse(!is.na(not_measurable), str_c("not_measurable: ", not_measurable), "no lsvi measurement")))) %>%
    filter(!reassessment) %>%
    filter(is.na(relocated_to)) %>%
    select(point_code, assessment_date, status, lsvi_measurement, relocation, reassessment, oversampling_subtype, type_target_all, type_observed, type_match, distance, remark)
```

