# Mhq admin data

```{r}
types <- read_types()
```


## Read google sheet data with mhq assessments admin

The INBO fieldwork coordinator uses [this googlesheet](https://docs.google.com/spreadsheets/d/1WkGPPYTb1Xytg5kJeOSxLvt3PUdFjlpPb4rwYG9znys/) to document which sampling units have been assessed.

To make the analysis reproducible we downloaded the 2021-04-26 version of the google sheet. 

```{r, eval = FALSE}
gs4_auth("toon.westra@inbo.be")

assessments_admin <- read_sheet(
    "https://docs.google.com/spreadsheets/d/1WkGPPYTb1Xytg5kJeOSxLvt3PUdFjlpPb4rwYG9znys/edit#gid=0",
    col_types = c("ccllcnnncDclncc"), trim_ws = TRUE
    )

```

```{r}
assessments_admin <- read_csv2("../../data/voortgang_mhq_terrestrisch_2021-04-26.csv") %>%
    select(-starts_with("...")) %>%
    mutate(legacy_site = ifelse(legacy_site == "WAAR", TRUE,
                                ifelse(legacy_site == "ONWAAR", FALSE, NA)),
           lsvi_measurement = ifelse(lsvi_measurement == "WAAR", TRUE,
                                ifelse(lsvi_measurement == "ONWAAR", FALSE, NA)),
           assessment_date = as.Date(assessment_date, format = "%d/%m/%Y"))
           
```


## Read standardized table of mhq refpoints, assessments and measurements

```{r}
mhq_terr_assessments <- read_vc("mhq_terr_assessments", 
                                root = "../../data")

mhq_terr_refpoints <- read_vc("mhq_terr_refpoints", 
                                root = "../../data")

mhq_terr_refpoints_check_update <- read_vc("mhq_terr_refpoints_check_update", 
                                root = "../../data")
    
mhq_terr_measurements <- read_vc("mhq_terr_measurements", 
                                root = "../../data")

mhq_terr_refpoints_validity <- read_vc("mhq_terr_refpoints_validity", 
                                root = "../../data")
```

# Update standardized tables

+ check for new assessments

```{r}
assessments_admin_done <- assessments_admin %>%
    filter(!is.na(assessment_date)) %>%
    rename(x_measured = x,
           y_measured = y)

assessments_table_todo <- mhq_terr_assessments %>%
    filter(is.na(assessment_date)) 

check_no_assessments <- all(is.na(assessments_table_todo$is_type_target)) &
     all(is.na(assessments_table_todo$assessment_source)) &
    all(is.na(assessments_table_todo$inaccessible))
    
assessments_table_new <- assessments_table_todo %>%
    select(point_code, type_target, x, y, fieldwork_team) %>%
    inner_join(assessments_admin_done, by = c("point_code", "fieldwork_team")) %>%
    mutate(distance_to_refpoint = sqrt((x_measured - x)^2 + (y_measured - y)^2),
           is_type_target = type_target == type_observed,
           is_type_target = ifelse(type_target == "2190" & type_observed == "2190_overig", TRUE, is_type_target),
           is_type_target = ifelse(is.na(is_type_target) & !lsvi_measurement, FALSE, is_type_target))

check_records <- assessments_table_new %>%
    filter(is.na(is_type_target) | is.na(lsvi_measurement))

```

Following records contain missing information:

```{r}
check_records %>%
    select(point_code, type_target, db_ref, assessment_date, assessment_source, lsvi_measurement, type_observed) %>%
    kbl() %>%
  kable_styling()
```
For following records the target type is observed but lsvi is not measured:

```{r}
check_records2 <- assessments_table_new %>%
  filter(is_type_target) %>%
  filter(!lsvi_measurement)

check_records2 %>%
    select(point_code, type_target, db_ref, assessment_date, assessment_source, lsvi_measurement, type_observed) %>%
    kbl() %>%
    kable_styling()
```


+ check for repeated assessments


```{r}
assessments_table_done <- mhq_terr_assessments %>%
    filter(!is.na(assessment_date))
  
assessments_admin_repeat <- assessments_admin_done %>%
    anti_join(assessments_table_new, by = c("point_code", "assessment_date")) %>%
    anti_join(assessments_table_done, by = c("point_code", "assessment_date"))

check_repeated_ass <- nrow(assessments_admin_repeat) > 0
  
```

+ check for anomalies

```{r}
assessments_table_done_check <- assessments_table_done %>%
  left_join(assessments_admin_done, by = c("point_code", "assessment_date"), suffix = c(".table", ".admin")) %>%
  mutate(match_assessment_source = assessment_source.table == assessment_source.admin,
         match_type_observed = type_observed.table == type_observed.admin)

check_type_observed <- assessments_table_done_check %>%
  filter(!match_type_observed)

check_missing_admin <- assessments_table_done_check %>%
  filter(is.na(lsvi_measurement))
```

+ check allow_other_types

```{r}
max_ranking <- mhq_terr_refpoints %>%
    filter(!is.na(grts_ranking_draw)) %>%
    group_by(type_target, sac) %>%
    summarise(max_grts_ranking_draw = max(grts_ranking_draw)) %>%
    ungroup() %>%
    rename(type_observed = type_target)

assessments_table_new <- assessments_table_new %>%
  mutate(type_observed = ifelse(type_observed == "2190_overig", "2190", type_observed)) %>%
  left_join(max_ranking, by = c("sac", "type_observed")) %>%
  group_by(point_code, assessment_date) %>%
  mutate(is_type_target_plot = sum(is_type_target) > 0) %>%
  ungroup() %>%
  mutate(allow_other_type = ifelse(is_type_target_plot, NA, (grts_ranking_draw <= max_grts_ranking_draw) & !is.na(max_grts_ranking_draw)))
  
check_allow_other_type <- assessments_table_new %>%
  filter(lsvi_measurement) %>%
  filter(!allow_other_type)
```

Following measurements are not valid because the observed type differs from the taget type and the grts-ranking is higher than the maximum grts-ranking in the sample of the observed type.

```{r}
check_allow_other_type %>%
  select(point_code, type_target, assessment_date, type_observed, lsvi_measurement, grts_ranking_draw, max_grts_ranking_draw, allow_other_type) %>%
  kbl() %>%
  kable_styling()
```



## Update mhq_terr_assessments

+ Create new sampling units when is_type_target is FALSE and a LSVI measurement is performed. For a limited number of sampling units allow_other_type is FALSE. Yet, for pragmatic reasons, we will keep these sampling units, altough theoretically they are not valid.

```{r}
# assessments_table_new <- assessments_table_new %>%
#   select(point_code, db_ref, type_target, is_type_target, type_observed, assessment_date, assessment_source, x = x_measured, y = y_measured, inaccessible, fieldwork_team, lsvi_measurement, allow_other_type)

new_sampling_units_strict <- assessments_table_new %>%
    filter(lsvi_measurement) %>%
    # filter(!is_type_target & !is_type_target_plot) %>%
    filter(allow_other_type)

new_sampling_units <- assessments_table_new %>%
    filter(lsvi_measurement) %>%
    filter(!is_type_target & !is_type_target_plot) %>%
    #filter(allow_other_type) %>%
    mutate(type_target = factor(type_observed, levels = levels(types$type)),
           is_type_target = TRUE,
           allow_other_type = NA) %>%
    unique()

assessments_table_new <- assessments_table_new %>%
  bind_rows(new_sampling_units)

mhq_terr_assessments_update <- mhq_terr_assessments %>%
  anti_join(assessments_table_new, by = c("point_code", "fieldwork_team")) %>%
  bind_rows(select(assessments_table_new, point_code, db_ref, type_target, is_type_target, type_observed, assessment_date, assessment_source, x = x_measured, y = y_measured, inaccessible, fieldwork_team))
  
```

To update the other standardized tables we will need the coordinates of the measured sampling units.For now we work with the available data to determine het number of measured valid sampling units per type. 

## Update mhq_terr_refpoints

```{r}
new_refpoints <- new_sampling_units %>%
  mutate(sampling_unit_code = str_c(grts_ranking_draw, "_", type_target)) %>%
  select(sampling_unit_code, type_target, point_code, sac, grts_ranking_draw, x = x_measured, y = y_measured, db_ref)

mhq_terr_refpoints_update <- mhq_terr_refpoints_check_update %>%
  bind_rows(new_refpoints)


```


