# Standing water

## Read admin data

The file `voortgang_mhq_watersurfaces_2025-11*25.csv` is in [this google drive](https://drive.google.com/drive/folders/19NsppwrC47ZAjv8Ndoh6aChkdud801T2).

It is the '2025-11-25' version of [this google sheet](https://docs.google.com/spreadsheets/d/1A0SUXxbs7Z4TB_lNPH11Wj0fTuFA9mfRslX-LzD4Rds/edit?gid=1191658174#gid=1191658174) in which the fieldwork progress of the 2nd monitoring cycle of mhq watersurfaces is documented.

```{r}

admin_cycle2_orig <- read_csv2("../../data/admin_mhq_watersurfaces/voortgang_mhq_watersurfaces_2025-11-25.csv")

admin_cycle2 <- admin_cycle2_orig %>%
    select(polygon_id = plygn_d, cycle1 = Meetnet_1, type_target = type_ll, screening = RednNM2, date_orig = `datum veldbezoek`, type_observed = `habtype veld`, lsvi_measurement = `LSVI-bepaling?`, reason_no_habitat = `reden geen habitat`, reason_not_measured = `geen opname wegens`, reason_not_measurable = `reden niet geschikt`, opmerking, in_inboveg =`ingevoerd in inboveg?`) %>%
    filter(!is.na(polygon_id)) %>%
    mutate(assessment_date = date_orig) %>%
    mutate(assessment_date = str_replace_all(assessment_date, "\\/", "\\-"),
           assessment_date = as.Date(assessment_date, format = "%d-%m-%Y"),
           lsvi_measurement = lsvi_measurement == "WAAR",
           in_inboveg = in_inboveg == "WAAR") %>%
    filter(!is.na(assessment_date) | !is.na(screening) | !is.na(reason_not_measurable)) %>%
    mutate(type_observed = ifelse(str_detect(screening, "gh") & !is.na(screening), "gh", type_observed),
           assessment_date = if_else(!is.na(screening) | is.na(assessment_date), as.Date("2024-01-30"), assessment_date)) %>% 
    group_by(polygon_id) %>%
    filter(assessment_date == max(assessment_date)) %>%
    ungroup() %>%
    mutate(reason_not_measured = ifelse(polygon_id == "LIMGNK0182", "tijdelijk ongeschikt", reason_not_measured),
           lsvi_measurement = ifelse(polygon_id == "LIMGNK0182", FALSE, lsvi_measurement)) #geen geschikt opname

write_csv2(admin_cycle2, "../../output/voortgang_mhq_aq_stilstaand_cycle2.csv")

```

## Update assessments

### Assessments cycle 2

-   Standardize type observed:
    -   separate row for each observed type
    -   check if observed is a valid type
    -   check if type observed is missing

```{r}

types <- read_types()

assessments_cycle2 <- admin_cycle2 %>%
    separate_longer_delim(cols = type_observed, delim = ",") %>%
    separate_longer_delim(cols = type_observed, delim =  "; ") %>%
    separate_longer_delim(cols = type_observed, delim =  "en ") %>%
    mutate(type_observed = str_trim(type_observed),
           is_valid_type = type_observed %in% c(as.character(types$type), "gh"))

assessments_cycle2_unsure <- assessments_cycle2 %>%
    filter(str_detect(type_observed, "\\?")) %>%
    select(polygon_id, assessment_date, type_observed, lsvi_measurement, opmerking)

# write_csv2(assessments_cycle2_unsure, "../../output/check_unsure.csv")
```

```{r}
assessments_cycle2 %>%
    filter(!is_valid_type) %>%
    select(polygon_id, screening, assessment_date, type_observed, reason_no_habitat, reason_not_measured, reason_not_measurable) %>%
    datatable(filter = "top",
              rownames = FALSE)
```

-   Change `3130n` to `3130_na` and `3130a` to `3130_aom`
-   Remove types ending with `?` (for example `3140?`): to be verified by field work coordinator

```{r}
assessments_cycle2 <- assessments_cycle2 %>%
    filter(!str_detect(type_observed, "\\?") | is.na(type_observed)) %>% 
    # filter((!reason_not_measurable %in% c("te hoge waterstand", "(grotendeels) drooggevallen")) | 
    #            is.na(reason_not_measurable)) %>%
    mutate(type_observed = ifelse(type_observed == "3130n", "3130_na",
                                  ifelse(type_observed == "3130a", "3130_aom", type_observed)),
           assessment_source = ifelse(str_detect(screening, "gh") & !is.na(screening), "orthophoto", "field assessment"),
           inaccessible = ifelse(reason_not_measurable == "geen toestemming om te betreden", "long term", NA),
           not_measurable = ifelse(reason_not_measured == "tijdelijk ongeschikt", "short term", NA),
           is_valid_type = type_observed %in% c(as.character(types$type), "gh"))
```

-   Create records for target types that are absent.
Two cases occur:
    -   `type_observed` = `gh`
        -   `is_present` is `FALSE` for all target types
        -   `no_habitat` is `TRUE`
    -   one or more types are observed but not all target types
        -   `is_present` is `FALSE` for the target types that are not observed
        -   `no_habitat` is `FALSE`

```{r}

assessments_cycle2_is_present <- assessments_cycle2 %>%
    filter(str_sub(type_observed, 1, 2) == "31" |
               type_observed == "2190_a") %>%
    mutate(is_present = TRUE,
           no_habitat = FALSE)

# in case a 31xx habitat is observed but another 31xx target habitat is not observed
assessments_cycle2_target_absent <- assessments_cycle2 %>%
    filter(str_sub(type_observed, 1, 2) == "31") %>%
    mutate(type_observed = type_target) %>%
    separate_longer_delim(cols = type_observed, delim =  "; ") %>%
    anti_join(assessments_cycle2_is_present, by = c("polygon_id", "type_observed")) %>%
    mutate(is_present = FALSE,
           no_habitat = FALSE,
           lsvi_measurement = FALSE) %>%
    unique()
    
assessments_cycle2_gh <- assessments_cycle2 %>%
    filter(type_observed == "gh") %>%
    separate_longer_delim(cols = type_target, delim =  "; ") %>%
    mutate(type_observed = type_target,
           is_present = FALSE,
           no_habitat = TRUE)

assessments_cycle2_unknown <- assessments_cycle2 %>%
    filter(is.na(type_observed)) %>%
    separate_longer_delim(cols = type_target, delim =  "; ") %>%
    mutate(type_observed = type_target,
           is_present = NA,
           no_habitat = NA,
           assessment_source = NA)

assessments_cycle2 <-  assessments_cycle2_is_present %>%
    bind_rows(assessments_cycle2_target_absent) %>%
    bind_rows(assessments_cycle2_gh) %>%
    bind_rows(assessments_cycle2_unknown) %>%
    arrange(assessment_date, polygon_id)

assessments_cycle2_stdz <- assessments_cycle2 %>%
    mutate(mhq_cycle = 2) %>%
    select(mhq_cycle, assessment_date, polygon_id, type = type_observed, is_present, no_habitat, assessment_source, inaccessible, not_measurable, reason_not_measurable)

```

Some checks:

-   Only one record per type and polygon_id combination?

```{r}
(assessments_cycle2_stdz %>%
    group_by(type, polygon_id) %>%
    filter(n() > 1) %>%
    nrow()) == 0
```
-   If `no_habitat` = `FALSE`, we should have at least one record with `is_present` = `TRUE`

```{r}
(assessments_cycle2_stdz %>%
    group_by(polygon_id) %>%
    mutate(no_habitat_check = !any(is_present)) %>%
    ungroup() %>%
    filter(no_habitat_check == no_habitat) %>%
     nrow()) == 0
```
-   All polygon_id's in the original table are persent in the standardized table

```{r}
all(admin_cycle2$polygon_id %in% assessments_cycle2_stdz$polygon_id)

```

### Assessments cycle 1

The previous versions of `mhq_watersurfaces_assessments`:

-   version 2023: all assessments in the first monitoring cycle
    -   generated with script `20_admin_standingwater.Rmd` in `habitatwatersurfaces_cycle1/src` folder of` n2khab-mhq-design` github repo
    -   note that watersurfaces have been visited for abiotic and biotic measurements in different years
    and that the assessment date corresponds with the first visit (usually for abiotic measurements)
-   version 2024: addition of rejected sampling units after desktop screening of selected sampling units
in the second monitoring cycle

```{r}

# alles van cyclus 1: assessment_date = minimum van datum biotische en datum abiotische opname
path_ws_c1 <- file.path(fileman_up("n2khab-mhq-design"), "habitatwatersurfaces_cycle1/output")

assessments_cycle1_orig <- read_vc(file = "mhq_watersurfaces_assessments", path_ws_c1)
measurements_cycle1 <- read_vc(file = "mhq_watersurfaces_measurements", path_ws_c1)
refpoints_v5 <- read_vc(file = "watersurfaces_refpoints", path_ws_c1) 

measurements_cycle1_biotic <- measurements_cycle1 %>%
    filter(measurement_type == "biotic") %>%
    distinct(polygon_id, measurement_date)

compare_date <- assessments_cycle1_orig %>%
    left_join(measurements_cycle1_biotic, by = "polygon_id")

path_ws_c2 <- file.path(fileman_up("n2khab-mhq-design"), "habitatwatersurfaces_cycle2/output")

# alles van cyclus 1 + desktop controle voor cyclus 2 van start ging
assessments_cycle1_bis <- read_vc(file = "mhq_watersurfaces_assessments", path_ws_c2)
popunits <- read_vc(file = "mhq_watersurfaces_populationunits", path_ws_c2) 

assessments_cycle1 <- assessments_cycle1_bis %>%
    group_by(polygon_id, assessment_date) %>%
    mutate(no_habitat = !any(is_present)) %>%
    ungroup() %>%
    mutate(mhq_cycle = ifelse(year(assessment_date) >= 2024, 2, 1))
```

### Combine

We add new assessments to create version 2025 of `mhq_watersurfaces_assessments`.

-   Are there any duplicate records?

```{r}
(assessments_cycle2 %>%
    semi_join(assessments_cycle1, by = c("polygon_id", "assessment_date")) %>%
    nrow()) > 0
```


```{r}
check_repeated <- assessments_cycle2_stdz %>%
    bind_rows(assessments_cycle1) %>%
    group_by(polygon_id) %>%
    filter(n_distinct(assessment_date) > 1) %>%
    ungroup() %>%
    arrange(polygon_id, assessment_date)

assessments_update <- assessments_cycle1 %>%
    bind_rows(assessments_cycle2_stdz) %>%
    arrange(assessment_date, polygon_id) %>%
    select(mhq_cycle, everything())
    
```

### Check

-   Check reference points:
    -   polygon_id `683951_v2014` has no reference point.
    It is not clear where it comes from as it is absent in watersurfaces_hab (v3 to v6) en the habitatmap
    -   we remove it from 'mhq_watersurfaces_assessments'
    
```{r}
mhq_watersurfaces_refpoints <- assessments_update %>%
    left_join(refpoints_v5, by = "polygon_id") 
 
missing_refpoint <- mhq_watersurfaces_refpoints %>%
    filter(is.na(grts_ranking))

assessments_update <- assessments_update %>%
    anti_join(missing_refpoint, by = "polygon_id")

```

```{r}
#nakijken
popunits_missing_polygon <- mhq_watersurfaces_refpoints %>%
    anti_join(popunits, by = "polygon_id") %>%
    filter(is_present & is.na(inaccessible))

```

### View result

The map shows:

-   assessments mhq cycle 1 (green circles)
-   assessments mhq cycle 2 (blue circles)

```{r}
mhq_watersurfaces_refpoints %>%
    filter(!is.na(x)) %>%
    mutate(show_color = ifelse(mhq_cycle == 1, "green", "blue")) %>%
    st_as_sf(coords = c("x", "y"), crs =31370) %>%
    st_transform(4326) %>%
    leaflet() %>%
    addTiles() %>%
    addCircleMarkers(color = ~show_color, label = ~str_c(type, ", is_present = ", is_present))
```

### Write results

Let's save the different versions in separate folders:

```{r}
data_path <- "../../data/mhq_watersurfaces"

if (!dir.exists(file.path(data_path, "v2023"))) {
    
    dir.create(file.path(data_path, "v2023"))
    
}

assessments_cycle1_orig %>%
    write_vc(file = "mhq_watersurfaces_assessments",
             root = file.path(data_path, "v2023"),
             sorting = c("assessment_date", "polygon_id", "type"),
             strict = FALSE)
```

```{r}
data_path <- "../../data/mhq_watersurfaces"

if (!dir.exists(file.path(data_path, "v2024"))) {
    
    dir.create(file.path(data_path, "v2024"))
    
}

assessments_cycle1_bis %>%
    write_vc(file = "mhq_watersurfaces_assessments",
             root = file.path(data_path, "v2024"),
             sorting = c("assessment_date", "polygon_id", "type"))
```


```{r}

if (!dir.exists(file.path(data_path, "v2025"))) {
    
    dir.create(file.path(data_path, "v2025"))
    
}

assessments_update %>%
    select(mhq_cycle, assessment_date, polygon_id, type, is_present, no_habitat, assessment_source, inaccessible, not_measurable, reason_not_measurable) %>%
    write_vc(file = "mhq_watersurfaces_assessments",
             root = file.path(data_path, "v2025"),
             sorting = c("mhq_cycle", "assessment_date", "polygon_id", "type"),
             strict = FALSE)
```

```{r}
check <- read_vc(file = "mhq_watersurfaces_assessments",
             root = file.path(data_path, "v2025"))
```


## Update measurements

The admin table also provides information on:

-   LSVI measurement
-   data imported in INBOVEG

To be checked: `lsvi_measurement = FALSE` `and in_inboveg = TRUE`

```{r}
assessments_cycle2 %>%
    filter(!lsvi_measurement & is_present & in_inboveg) %>%
    select(polygon_id, assessment_date, type_observed, lsvi_measurement, in_inboveg) %>%
    datatable(rownames = FALSE)
```

```{r}
measurements_cycle2 <- assessments_cycle2 %>%
    mutate(lsvi_measurement = ifelse(polygon_id == "ANTTUR0296" & type_observed == "3160", FALSE, lsvi_measurement)) %>% #based on opmerking
    mutate(lsvi_measurement = ifelse(polygon_id == "WVLHAR0329", TRUE, lsvi_measurement)) %>% #na controle
    filter(lsvi_measurement) %>%
    filter(is_present) %>%
    filter(is.na(reason_not_measured)) %>%
    select(polygon_id, measurement_date = assessment_date, type = type_observed, is_present, lsvi_measurement, reason_not_measured, reason_not_measurable, opmerking, inaccessible, not_measurable, in_inboveg) %>%
    mutate(measurement_year = year(measurement_date))
```

### Check the measurements in INBOVEG

We check which measurements are imported in INBOVEG.
The `git2rdata` file `HT31xx_header` in the [n2khqb-mhq_data](https://github.com/inbo/n2khab-mhq-data/tree/master/processed/inboveg_mhq_aq) repo gives an overview of the mhq watersurfaces measurements imported in INBOVEG.

```{r}
path_inboveg <- file.path(fileman_up("n2khab-mhq-data"), "processed/inboveg_mhq_aq")

ws_inboveg <- read_vc("HT31xx_header", 
                      root = path_inboveg)

ws_codes <- read_csv2(file.path(path_inboveg, "watervlakken_codes/mapping_table_inbocode.csv")) %>%
    select(polygon_id = codeplas, inbocode)

ws_inboveg <- ws_inboveg %>%
    select(location, type = type_observed, measurement_date = date, reliability, suitable_mhq, measured, recording_givid, user_reference) %>%
    left_join(ws_codes %>%
                  rename(location = inbocode),
              by = "location") %>%
    mutate(polygon_id = ifelse(is.na(polygon_id), location, polygon_id)) %>%
    mutate(measurement_date = as.Date(measurement_date),
           measurement_year = year(measurement_date)) 
```

### Update measurements cycle 1

We update mhq_watersurfaces_measurements for cycle 1:

-   add recording_givid of INBOVEG
-   add exact measurement date
-   only select biotic measurements
-   only select measurements that are imported in INBOVEG
-   only select measurements with suitable_MHQ = TRUE

```{r}

measurements_cycle1_biotic <- measurements_cycle1 %>%
    group_by(polygon_id) %>%
    mutate(assessment_date = min(measurement_date),
           measurement_year = year(measurement_date)) %>%
    ungroup() %>%
    filter(measurement_type == "biotic")

# type is missing in some cases; this is the case when the target type is not present; to be removed
check_missing_type <- measurements_cycle1_biotic %>%
    filter(is.na(type)) %>%
    select(-type) %>%
    left_join(assessments_cycle1,
              by = c("assessment_date", "polygon_id"))
    
check_presence <- measurements_cycle1_biotic %>%
    left_join(assessments_cycle1 %>%
                  select(assessment_date, polygon_id, type, is_present),
              by = c("assessment_date", "polygon_id", "type"))

measurements_cycle1_inboveg <- measurements_cycle1_biotic %>%
    left_join(assessments_cycle1 %>%
                  select(assessment_date, polygon_id, type, is_present),
              by = c("assessment_date", "polygon_id", "type")) %>%
    left_join(ws_inboveg,
              by = c("polygon_id", "type", "measurement_year"),
              suffix  = c("_admin", "_inboveg")
                  )

measurements_cycle1_match <- measurements_cycle1_inboveg %>%
    filter(!is.na(recording_givid)) %>%
    filter(suitable_mhq) %>%
    mutate(mhq_cycle = 1)

```

These are the resulting measurements in cycle 1 for which the measurement year in the admin table matches the measurement year in INBOVEG.

```{r}
measurements_cycle1_match %>%
    select(mhq_cycle, polygon_id, measurement_date = measurement_date_inboveg, type, user_reference, recording_givid, suitable_mhq) %>%
    datatable(rownames = FALSE,
              filter = "top")
```

These are the measurements in cycle 1 for which the measurement year in the admin table does not matches the measurement year in INBOVEG.
We consider the measurement date in INBOVEG as the correct one.

```{r}
check_mismatch_year <- measurements_cycle1_biotic %>%
    anti_join(measurements_cycle1_match, by = c("polygon_id", "type", "assessment_date")) %>%
    left_join(assessments_cycle1 %>%
                  select(assessment_date, polygon_id, type, is_present),
              by = c("assessment_date", "polygon_id", "type")) %>%
    left_join(ws_inboveg,
              by = c("polygon_id", "type"),
              suffix = c("_admin", "_inboveg")
                  ) %>%
    filter(!is.na(recording_givid)) %>%
    filter(suitable_mhq) %>%
    filter(measurement_year_inboveg <= 2022) %>%
    mutate(diff_years = abs(measurement_year_admin - measurement_year_inboveg))

measurements_cycle1_match_corr <- check_mismatch_year %>%
    mutate(mhq_cycle = 1) 
    
check_mismatch_year %>%
    select(assessment_date, polygon_id, type, is_present, measurement_year_admin, measurement_year_inboveg, diff_years, measurement_date_inboveg) %>%
    arrange(diff_years) %>%
    datatable(rownames = FALSE,
              filter = "top")
```

For the following measurements no matching INBOVEG record is found
When `loc_in_inboveg` is `TRUE` there is a measurement in INBOVEG for the `polyogn_id`.
This occurs in two occasions:

-   there is a measurement in INBOVEG for another type (specified in `types_inboveg`)
-   `suitable_mhq` is `FALSE`

```{r}

check_missing_inboveg <-  measurements_cycle1_inboveg %>%
    filter(!is.na(type)) %>%
    filter(is.na(recording_givid)) %>%
    anti_join(measurements_cycle1_match, by = c("assessment_date", "polygon_id", "type")) %>%
    anti_join(measurements_cycle1_match_corr, by = c("assessment_date", "polygon_id", "type")) %>%
    left_join(popunits, by = c("polygon_id", "type")) %>%
    select(-suitable_mhq) %>%
    mutate(loc_in_inboveg = polygon_id %in% (ws_inboveg %>%
                  filter(measurement_year <= 2022))$polygon_id) %>% 
    left_join(ws_inboveg %>%
                  filter(measurement_year <= 2022) %>%
                  group_by(polygon_id, suitable_mhq) %>%
                  summarise(types_inboveg = str_c(unique(type), collapse = ";")),
              by = c("polygon_id")) %>% 
    select(polygon_id, db_ref, type, measurement_year, assessment_date, loc_in_inboveg, types_inboveg, suitable_mhq)

check_missing_inboveg %>%
    datatable(rownames = FALSE,
              filter = "top")
```

```{r}
measurements_cycle1 <- measurements_cycle1_match %>%
    bind_rows(measurements_cycle1_match_corr) %>%
    select(mhq_cycle, polygon_id, assessment_date, measurement_date = measurement_date_inboveg, type, user_reference, recording_givid, suitable_mhq)
```

### Measurements cycle2

We first check for which measurements we find a record in INBOVEG for the same `polygon_id`, `type` and `measurement_date`.

```{r}
measurements_cycle2_match <- measurements_cycle2 %>%
    inner_join(ws_inboveg, by = c("polygon_id", "type", "measurement_date")) %>%
    filter(suitable_mhq) %>%
    mutate(mhq_cycle = 2)
```

```{r}
measurements_cycle2_match %>%
    select(polygon_id, measurement_date, type, suitable_mhq, recording_givid) %>%
    datatable(rownames = FALSE,
              filter = "top") 
```
Next we check mismatches in `measurement_date`.

```{r}
check_date <- measurements_cycle2 %>%
    anti_join(measurements_cycle2_match, by = c("polygon_id", "type", "measurement_date")) %>%
    inner_join(ws_inboveg, by = c("polygon_id", "type", "measurement_year"), suffix = c("_admin", "_inboveg")) %>%
    select(polygon_id, type, measurement_year, starts_with("measurement_date"))

check_date %>%
    datatable(rownames = FALSE)
```

```{r}
check_type <- measurements_cycle2 %>%
    anti_join(measurements_cycle2_match, by = c("polygon_id", "type", "measurement_date")) %>%
    inner_join(ws_inboveg, by = c("polygon_id", "measurement_date"), suffix = c("_admin", "_inboveg")) %>%
    select(polygon_id, starts_with("type"), starts_with("measurement_date"))
```

Following table shows measurements for which the observed habitat type in the admin table is `gh`, 
yet in INBOVEG it is recorded that the target type is present. 

```{r}
check_is_present <- assessments_cycle2_stdz %>%
    filter(!is_present) %>%
    left_join(ws_inboveg, by = c("polygon_id", "type")) %>%
    filter(year(assessment_date) == year(measurement_date))

check_is_present %>%
    select(assessment_date, polygon_id, type, is_present, measurement_date, suitable_mhq, measured) %>%
    datatable(rownames = FALSE)
    
```

Following table gives an overview of the measurements not imported in INBOVEG yet.

```{r}
check_missing_inboveg <- measurements_cycle2 %>%
    anti_join(measurements_cycle2_match, by = c("polygon_id", "type", "measurement_date")) %>%
    anti_join(check_date, by =  c("polygon_id", "type", "measurement_year"))

check_missing_inboveg %>%
    select(polygon_id, type, measurement_date, in_inboveg) %>%
    datatable(rownames = FALSE,
              filter = "top")
```
There are also measurements in INBOVEG that are not mentioned in the admin table (combination of `measurement_date`, `polygon_id`, `type` does not occur).

We first check if the watersurface is included in the sample of the second monitoring cycle.
For this we use the gpkg `mhq_standingwater_cycle2_2024-04-17.gpkg` which can be found in [this google drive folder](https://drive.google.com/drive/folders/1Nz6_QApHBdeg4oBmUlAcsgoEQQe3m9xI).

We also determine the strata (`in_sac` and `area_class`) of the watersurfaces, as in some strata we measure all watersurfaces.
All measurements in these strata are relevant for mhq.
We determine the strata based on [version 2024 of the `watersurfaces` data source](https://zenodo.org/records/14203168).

```{r}
mhq_watersurfaces_cycle2_sf <- st_read("../../data/design_cycle2/mhq_standingwater_cycle2_2024-04-17.gpkg")

mhq_watersurfaces_cycle2 <- mhq_watersurfaces_cycle2_sf %>%
    st_drop_geometry() %>%
    # separate_longer_delim(cols = type_all, delim = "; ") %>%
    select(polygon_id, grts_ranking, grts_ranking_draw, type_all, area_class, sac, panel)

watersurfaces <- read_watersurfaces()
watersurfaces_hab <- read_watersurfaces_hab()
sac <- read_admin_areas(dsn = "sac")

ws_area_class <- watersurfaces %>%
    filter(polygon_id %in% ws_inboveg$polygon_id) %>%
    st_join(sac %>%
                select(sac_code), 
            largest = TRUE) %>%
    mutate(in_sac = !is.na(sac_code),
           area_ha = drop_units(st_area(geom) / 10000),
           area_class = ifelse(area_ha <= 1, "area <= 1 ha",
                               ifelse(area_ha <= 5, "1 ha < area <= 5 ha",
                                      ifelse(area_ha <= 50, "5 ha < area < 50 ha",
                                             "area >= 50 ha")))) %>%
    st_drop_geometry() %>%
    select(polygon_id, area_class, in_sac)

wsh_types <- watersurfaces_hab$watersurfaces_types %>%
    group_by(polygon_id) %>%
    summarise(type_wsh = str_c(type, collapse = "; "))

check_missing_admin <- ws_inboveg %>%
    anti_join(measurements_cycle2_match, by = c("polygon_id", "type", "measurement_date")) %>%
    anti_join(measurements_cycle1, by = c("polygon_id", "type", "measurement_date")) %>%
    anti_join(check_date, by =  c("polygon_id", "type", "measurement_year")) %>%
    # filter(measurement_year >= 2022) %>%
    filter(suitable_mhq) %>%
    left_join(ws_area_class,
              by = "polygon_id") %>%
    left_join(wsh_types,
              by = "polygon_id") %>%
    left_join(mhq_watersurfaces_cycle2 %>%
                  select(polygon_id, type_target = type_all, panel),
              by = c("polygon_id")) %>%
    mutate(loc_in_admin = polygon_id %in% admin_cycle2_orig$plygn_d,
          ws_in_c2 = !is.na(type_target),
           ws_integraal = type == "3110" | 
               area_class == "area >= 50 ha" | 
               (area_class == "1 ha < area <= 5 ha" & type == "2190_a"),
           type_target_in_c2 = str_detect(type_target, as.character(type)))

```

For following measurements, the watersurfaces are included in the sample of mhq cycle 2.

```{r}
check_missing_admin %>%
    filter(ws_in_c2) %>%
    arrange(measurement_date) %>%
    select(polygon_id, location, type_observed = type, measurement_date, recording_givid, suitable_mhq, type_target) %>%
    filter(str_detect(type_target, as.character(type_observed))) %>%
    arrange(desc(measurement_date)) %>%
    datatable(rownames = FALSE,
              filter = "top")
```

Some additional measurements belong to strata which are fully measured (type 3110 and watersurfaces >= 50 ha):

```{r}

check_missing_admin %>%
    filter(!ws_in_c2) %>%
    filter(ws_integraal) %>%
    select(polygon_id, location, area_class, type_observed = type, measurement_date, recording_givid, suitable_mhq) %>%
    datatable(rownames = FALSE,
              filter = "top")

```

All other measurements:

```{r}
check_missing_admin %>%
    filter(!ws_in_c2) %>%
    filter(!ws_integraal) %>%
    arrange(desc(measurement_date)) %>%
    select(polygon_id, location, area_class, type_observed = type, measurement_date, recording_givid, suitable_mhq) %>%
    datatable(rownames = FALSE,
              filter = "top")
```

### Update assessments table based on measurements table

#### Correct assessment date based on measurement date

The admin table in cycle 1 only contains the year and not the exact date of the assessments.
Therefore the assessment date was set as year-01-01 
INBOVEG does contain the exact measurement date.
We correct the assessment date based on the measurement date (for those watersurfaces that were assessed and measured).

```{r}
assessments_update_corr_date <- assessments_update %>%
    inner_join(measurements_cycle1_match %>%
                   filter(year(assessment_date) == measurement_year) %>%
                  select(assessment_date, measurement_date_inboveg, type, polygon_id),
              by = c("polygon_id", "assessment_date", "type")) 

assessments_update <- assessments_update %>%
    anti_join(assessments_update_corr_date, by = c("polygon_id", "assessment_date", "type")) %>%
    bind_rows(assessments_update_corr_date %>%
                  mutate(assessment_date = measurement_date_inboveg) %>%
                  select(-measurement_date_inboveg))
```

#### Additional assessments based on measurements

The admin table of cycle1 conains per location the date of the biotic and abiotic measurments.
The minimum of both dates was added to the assessment table.
In general this was the date of the abiotic measurement.

However both dates are relevant, as often there are several years between both measurement types.
We add the date of the measurement that was not included in the previous version of the assessment table, and consider this as an additional assessment.

```{r}
new_assessments <- measurements_cycle1 %>%
    filter(year(assessment_date) != year(measurement_date)) %>%
    select(polygon_id, type, assessment_date = measurement_date) %>%
    mutate(is_present = TRUE,
           no_habitat = FALSE,
           assessment_source = "field assessment",
           mhq_cycle = 1)  %>%
    anti_join(assessments_update, by = c("polygon_id", "type", "assessment_date"))

# check if there are new assessments in the same year as exisiting assessments
check <- new_assessments %>%
    mutate(assessment_year = year(assessment_date)) %>%
    semi_join(assessments_update %>%
                  mutate(assessment_year = year(assessment_date)),
              by = c("polygon_id", "type", "assessment_year"))

assessments_update <- assessments_update %>%
    bind_rows(new_assessments)
```

We also add assessments that are missing in the admin table.

```{r}
mhq_assessmnets_add_presence <- check_missing_admin %>%
    select(polygon_id, type, assessment_date = measurement_date) %>%
    mutate(is_present = TRUE,
           no_habitat = FALSE,
           assessment_source = "field assessment",
           mhq_cycle = ifelse(year(assessment_date) < 2022, 1, 2)) %>%
    anti_join(assessments_update, by = c("polygon_id", "type", "assessment_date"))

check <- check_missing_admin %>%
    mutate(assessment_date = measurement_date) %>%
    inner_join(assessments_update, by =  c("polygon_id", "type", "assessment_date"))

assessments_update2 <- assessments_update %>%
    select(mhq_cycle, assessment_date, polygon_id, type, is_present, no_habitat, assessment_source, inaccessible, not_measurable, reason_not_measurable) %>%
    bind_rows(mhq_assessmnets_add_presence) %>%
    arrange(mhq_cycle, assessment_date, polygon_id, type)

check <- assessments_update2 %>%
    group_by(polygon_id, assessment_date, type) %>%
    filter(n() > 1)

# if two assessments in the same year: remove the assessment in january (default date, when date is unknown) 
remove_assessm <- assessments_update2 %>%
    mutate(assessment_year = year(assessment_date)) %>%
    group_by(mhq_cycle, polygon_id, type, assessment_year) %>%
    filter(n() > 1 & month(assessment_date) == 1) %>%
    ungroup()

assessments_update2 <- assessments_update2 %>%
    anti_join(remove_assessm, by = c("polygon_id", "type", "assessment_date"))

assessments_update2 %>%
    write_vc(file = "mhq_watersurfaces_assessments",
             root = file.path(data_path, "v2025"),
             sorting = c("mhq_cycle", "assessment_date", "polygon_id", "type"),
             strict = FALSE)

assessments_update2 %>%
    write_vc(file = "mhq_watersurfaces_assessments",
             root = data_path,
             sorting = c("mhq_cycle", "assessment_date", "polygon_id", "type"),
             strict = FALSE)

```

### Standardize measurement table

```{r}

duplicate <- measurements_cycle1 %>% 
    semi_join(measurements_cycle2_match, by = c("polygon_id", "measurement_date", "type"))

measurements_update <- measurements_cycle1 %>%
    anti_join(measurements_cycle2_match, by = c("polygon_id", "measurement_date", "type")) %>%
    bind_rows(measurements_cycle2_match) %>%
    bind_rows(check_missing_admin %>%
                 mutate(mhq_cycle = ifelse(year(measurement_date) < 2022, 1, 2))) %>%
    select(mhq_cycle, measurement_date, polygon_id, type, user_reference, recording_givid) %>%
    arrange(measurement_date, polygon_id, type)

check <- measurements_update %>%
    group_by(polygon_id, measurement_date, type) %>%
    filter(n() > 1)

measurements_update  %>%
    select(mhq_cycle, measurement_date, polygon_id, type, user_reference, recording_givid) %>%
    write_vc(file = "mhq_watersurfaces_measurements",
             root = data_path,
             sorting = c("measurement_date", "polygon_id", "type"),
             strict = FALSE)

measurements_update  %>%
    select(mhq_cycle, measurement_date, polygon_id, type, user_reference, recording_givid) %>%
    write_vc(file = "mhq_watersurfaces_measurements",
             root = file.path(data_path, "v2025"),
             sorting = c("measurement_date", "polygon_id", "type"),
             strict = FALSE)

```
## Overview checksums

```{r}

hashes_data <-
    tibble(filepath = str_c(data_path, "/",
        list.files(path = data_path,
            recursive = TRUE)
      )) %>%
    mutate(filename = str_match(filepath, "(.+\\/)*(.+)")[,3],
           md5 = map(filepath, function(x) {
                           file(x) %>% md5 %>% str_c(collapse = '')
                         }) %>% as.character) %>%
    filter(str_detect(filename, "measure|assess")) %>%
    mutate(filepath = str_remove(filepath, "../../"),
           folder = str_remove(filepath, filename)) %>%
    select(folder, filename, md5)

hashes_data %>%
    kbl()
```


