# Streams

## Progress cycle 2

The second monitoring cycle of the 3260 mhq scheme started in 2022 and last until 2027.

It appears that many of the selected sampling units do not contain 3260 habitat. Therefore we have to select new sampling units to achieve a sufficient sample size.

## Selected sampling units

The file 'meetpunten3260_cyclus2_versie2022-05-05.csv' list the selected sampling units for cycle 2.
The file can be found in this [google drive folder](https://drive.google.com/drive/folders/1UNKCLEgiYq2r6EkfWXMkppSDraPcqfAU).

```{r}
sample_cycle2 <- read_csv2("../../data/design_cycle2/meetpunten3260_cyclus2_versie2022-05-05.csv")

path_design <- fileman_up("n2khab-mhq-design")

samplingframe_habitatstreams_1_7 <- read_vc(file = "samplingframe_habitatstreams_v1_7", 
                                            root = file.path(path_design, "habitatstreams_cycle2/data"))
```

```{r}
sample_cycle2 %>%
  group_by(sac_streams) %>%
  summarize(sample_size = n()) %>%
  ungroup() %>%
  kable() %>%
  kable_styling()
```

```{r, eval = FALSE}
admin_streams_cycle2_orig <- read_csv("../../data/admin_mhq_watersurfaces/voortgang_mhq_aq_stromend_2025-03-17.csv")

assessments_cycle2 <- admin_streams_cycle2_orig %>%
    select(segment_id = sampling_unit_code, assessment_date = `datum veldbezoek`, type = doelhabitat) %>%
    filter(!is.na(segment_id), !is.na(type))

write_csv2(assessments_cycle2, "../../output/voortgang_mhq_aq_stromend_cycle2.csv")
```

## Visited samplign units

The file 'Monitoring3260_SVZ2024_voorToonWestra.csv' contains an overview of the visited and measured sampling units.


```{r}
admin_streams_orig <- read_csv2("../../data/admin_mhq_watersurfaces/Monitoring3260_SVZ2024_voorToonWestra.csv")

admin_streams <- admin_streams_orig %>%
  mutate(assessment_year = `jaar monitoring`,
         assessment_date = as.Date(`datum veldbezoek`, format = "%d/%m/%Y"),
         sac_streams = ifelse(SBZ == 1, "intersecting or downstream",
                      ifelse(SBZ == 0, "outside", SBZ)),
         type_observed = doelhabitat,
         assessment_date = if_else(is.na(assessment_date), as.Date(str_c(assessment_year, "-01-01")), assessment_date),
         measured = klasse == "mon3260") %>%
  select(field_code = Veldcode, segment_id = sampling_unit_code,assessment_year, assessment_date, type_observed, klasse, measured, reason_no_assessment = `reden geen bemonstering/ongeschikt`, x_assessment = x, y_assessment = y, position_change = `locatie SO of SA verschoven (xy aangepast tov origineel aangereikt meetpunt, volgens protocol)`) %>%
    filter(!is.na(field_code))

admin_cycle2 <- admin_streams %>%
  filter(assessment_year >= 2022) %>%
  group_by(field_code) %>%
  filter(assessment_date == max(assessment_date)) %>%
  ungroup() 

# replacement
admin_cycle2 <- admin_cycle2 %>%
  mutate(segment_id = ifelse(field_code == "KaaBe_01a", "1_4_772", segment_id),
         klasse = ifelse(field_code == "KaaBe_01", "uitval (ongeschikt)", klasse),
         type_observed = ifelse(field_code == "AntGr_01", "3260", type_observed))
```


## Assessments in cycle 2


```{r}
mhq_streams_refpoints <- read_vc(root =  "../../data/mhq_streams/v2022", file = "mhq_streams_refpoints")

assessments_cycle2 <- sample_cycle2 %>%
    select(-field_code) %>%
    select(segment_id = sampling_unit_code, position_change_c1 = punt_verplaatst, grts_ranking_draw, type_meetpunt, x, y) %>%
    left_join(admin_cycle2, by = "segment_id") %>%
    mutate(klasse = ifelse(is.na(klasse), "nog niet bemonsterd", klasse),
           point_code = str_c(grts_ranking_draw, ifelse(position_change_c1 == "ja" & !is.na(position_change_c1),"_2", "_1"))) %>%
    left_join(select(mhq_streams_refpoints, segment_id, point_code, grts_ranking), by = c("point_code", "segment_id")) %>%
    mutate(distance = round(sqrt((x - x_assessment) ^ 2 + (y - y_assessment) ^ 2), 1))

# point codes for new sampling units not in mhq_streams_refpoints yet
assessments_cycle2 <- assessments_cycle2 %>%
    mutate(grts_ranking = ifelse(type_meetpunt == "Nieuw meetpunt", grts_ranking_draw, grts_ranking))

# point codes for segment_id's 1497974_1, 1497974_2, 131970_1 and 131970_2 missing in mhq_streams_refpoints
grts_master <- read_GRTSmh()

grts_ranking_missing <- assessments_cycle2 %>%
    filter(is.na(grts_ranking)) %>%
    st_as_sf(coords = c("x", "y"), crs = 31370, remove = FALSE) 

grts_ranking_missing <- grts_ranking_missing %>%
    mutate(grts_ranking = (terra::extract(grts_master, grts_ranking_missing))$GRTSmaster_habitats) %>%
    st_drop_geometry()

# position change in cycle 2: modify point_code and grts_ranking

su_position_change <- assessments_cycle2 %>%
    filter(position_change == "ja") %>%
    mutate(point_code = str_c(grts_ranking_draw, "_2")) %>%
    st_as_sf(coords = c("x_assessment", "y_assessment"), crs = 31370, remove = FALSE) 

su_position_change <- su_position_change %>%
    mutate(grts_ranking = (terra::extract(grts_master, su_position_change))$GRTSmaster_habitats) %>%
    st_drop_geometry()
    
assessments_cycle2 <- assessments_cycle2 %>%
    anti_join(grts_ranking_missing, by = "point_code") %>%
    bind_rows(grts_ranking_missing) %>%
    anti_join(su_position_change, by = c("grts_ranking_draw", "position_change")) %>%
    bind_rows(su_position_change)
```




## Update mhq_streams_assessments

```{r}
mhq_assessments_v2020 <- read_vc(root =  "../../data/mhq_streams/v2020", file = "mhq_streams_assessments")
mhq_streams_refpoints_v2020 <- read_vc(root =  "../../data/mhq_streams/v2020", file = "mhq_streams_refpoints")
```

```{r}
mhq_assessments_v2022 <- read_vc(root =  "../../data/mhq_streams/v2022", file = "mhq_streams_assessments")

mhq_streams_refpoints_validity <- read_vc(root =  "../../data/mhq_streams/v2022", file = "mhq_streams_refpoints_validity")

check_missing <- mhq_assessments_v2020 %>%
    anti_join(mhq_assessments_v2022, by = c("segment_id", "point_code", "assessment_date")) %>%
    filter(!is.na(assessment_date)) 

check_missing_2020 <- admin_streams %>%
    anti_join(mhq_assessments_v2020, by = "segment_id") %>%
    filter(!is.na(segment_id)) %>%
    filter(year(assessment_date) < 2020)

mhq_streams_assessments_new <- assessments_cycle2 %>%
  filter(!is.na(assessment_date)) %>%
  mutate(type_observed = ifelse(type_observed %in% c("gh", "3260"), type_observed, NA),
         inaccessible = ifelse(klasse == "uitval (ontoegankelijk)", "long term", NA),
         not_measurable = ifelse(klasse == "uitval (ongeschikt)", "long term",
                                 ifelse(klasse == "uitval (tijdelijk ongeschikt)", "short term", NA)),
         change_location = position_change == "ja",
         assessment_source = "field assessment") %>%
  select(point_code, segment_id, grts_ranking, grts_ranking_draw, assessment_date, type_observed, measured, klasse, inaccessible, not_measurable, reason_no_assessment, x = x_assessment, y = y_assessment, x_orig = x, y_orig = y, change_location, assessment_source)

check3 <- mhq_streams_assessments_new %>%
    group_by(segment_id) %>%
    filter(n() > 1)

types <- read_types()

reason_cycle1 <- admin_streams %>%
    filter(year(assessment_date) < 2022) %>%
    filter(klasse != "niet meetellen (2x bezocht)") %>%
    select(assessment_date, segment_id, reason_no_assessment, measured, klasse) 

check <- reason_cycle1 %>%
    group_by(assessment_date, segment_id) %>%
    filter(n() > 1)

mq_streams_assessments_cycle1 <- mhq_assessments_v2022 %>%
    bind_rows(check_missing) %>%
  group_by(segment_id) %>%
    mutate(replacement = n_distinct(point_code) > 1) %>%
    ungroup() %>%
  mutate(change_location = str_detect(point_code, "_2"),
         type_observed = ifelse(segment_id %in% c("1_7_170694", "1_7_7841"), "gh", type_observed), #checked in admin table
         assessment_source = ifelse(segment_id %in% c("1_7_170694", "1_7_7841"), "orthophoto", assessment_source)) %>% #checked in admin table
  filter(!is.na(assessment_date))

mq_streams_assessments_cycle1_pos1 <- mq_streams_assessments_cycle1 %>%
    filter(str_detect(point_code, "_1")) %>%
    left_join(reason_cycle1, by = c("segment_id", "assessment_date")) %>%
    mutate(inaccessible = ifelse(str_detect(reason_no_assessment, "permanent niet toegankelijk"), "long term",
                                 ifelse(str_detect(reason_no_assessment, "tijdelijk niet toegankelijk"), "short term", NA)),
           not_measurable = ifelse(str_detect(reason_no_assessment, "permanent ongeschikt") & is.na(inaccessible), "long term",
                                 ifelse(str_detect(reason_no_assessment, "tijdelijk niet geschikt|tijdelijk ongeschikt") & is.na(inaccessible), "short term", NA))) %>%
    mutate(not_measurable = ifelse(replacement, "long term", not_measurable), #originele positie van verplaatste punten
           measured = ifelse(replacement, FALSE, measured),
           klasse = ifelse(replacement, "uitval (ongeschikt)", klasse)
           )

mq_streams_assessments_cycle1_pos2 <- mq_streams_assessments_cycle1 %>%
    filter(str_detect(point_code, "_2")) %>%
    left_join(reason_cycle1, by = c("segment_id", "assessment_date")) %>%
    mutate(reason_no_assessment = ifelse(segment_id == "1_4_1497", 
                                         "permanent ongeschikt: schaduw en bodemplaat", reason_no_assessment),  #manual check
           measurable = ifelse(segment_id == "1_4_838", FALSE, measurable),
           measured = measurable & type_observed == "3260",
           not_measurable = ifelse(measurable, NA, "long term"),
           reason_no_assessment = ifelse(measurable, NA, reason_no_assessment),
           klasse = ifelse(measurable,
                           ifelse(type_observed == "3260", "mon3260", "uitval (gh)"),
                           "uitval (ongeschikt)"))

mq_streams_assessments_update <- mq_streams_assessments_cycle1_pos1 %>%
    bind_rows(mq_streams_assessments_cycle1_pos2) %>%
  bind_rows(mhq_streams_assessments_new) %>%
  mutate(type = "3260",
         type = factor(type, levels = levels(types$type)),
         is_present = type_observed == type,
         no_habitat = str_detect(type_observed, "gh")) 

mq_streams_assessments_output <- mq_streams_assessments_update %>%
  select(assessment_date, point_code, segment_id, type, is_present, no_habitat, assessment_source, measured, inaccessible, not_measurable, reason_no_assessment, change_location) %>%
    arrange(assessment_date)

check <- mq_streams_assessments_output %>%
    group_by(segment_id, point_code, assessment_date) %>%
    filter(n() > 1)

```

### Check results

```{r}
check_replaced <- mq_streams_assessments_output %>%
    mutate(cycle = ifelse(year(assessment_date) < 2022, 1, 2)) %>%
    filter(str_detect(point_code, "_2")) %>%
    group_by(point_code) %>%
    mutate(n_cycles = n_distinct(cycle)) %>%
    ungroup()

```


```{r}
mq_streams_assessments_output %>%
    group_by(type, is_present, measured, inaccessible, not_measurable) %>%
    summarise(n_rec = n()) %>%
    kable() %>%
    kable_styling()
```

### write results

```{r}
mq_streams_assessments_output %>%
    write_vc(file = "mhq_streams_assessments",
             root = "../../data/mhq_streams",
             sorting = c("assessment_date", "point_code"))
```


## Update mhq_streams_refpoints

```{r}
check_missing_refpoint_2022 <- mhq_streams_refpoints_v2020 %>%
    anti_join(mhq_streams_refpoints, by = c("point_code")) %>%
    semi_join(mq_streams_assessments_update, by = c("point_code"))

mhq_streams_refpoints <- mhq_streams_refpoints%>%
    bind_rows(check_missing_refpoint_2022)

check <- mhq_streams_refpoints %>%
    group_by(point_code) %>%
    filter(n() > 1)

mhq_streams_refpoints_new <- mq_streams_assessments_update %>%
    anti_join(mhq_streams_refpoints, by = "point_code")

mhq_streams_refpoints_update <- mhq_streams_refpoints %>%
    bind_rows(mhq_streams_refpoints_new) %>%
    select(point_code, segment_id, grts_ranking, x, y)

check <- mq_streams_assessments_output %>%
    anti_join(mhq_streams_refpoints_update, by = "point_code")
```

```{r}
mhq_streams_refpoints_update %>%
    st_as_sf(coords = c("x", "y"), crs = 31370) %>%
    st_transform(4326) %>%
    leaflet() %>%
    addTiles() %>%
    addCircleMarkers()
```

```{r}
mhq_streams_refpoints_update %>%
    write_vc(file = "mhq_streams_refpoints",
             digits = 20,
             root = "../../data/mhq_streams",
             sorting = c("grts_ranking", "point_code"))
```


## Update mhq_streams_popunits

```{r}
mhq_streams_popunits <-  mhq_streams_refpoints %>%
    bind_rows(mhq_streams_refpoints_new) %>%
    select(point_code, grts_ranking, grts_ranking_draw, sac_streams, legacy_site, type, segment_id, db_ref) %>%
    left_join(select(mq_streams_assessments_update, point_code, assessment_date, measured, inaccessible, not_measurable, is_present), by = "point_code") %>%
    group_by(point_code) %>%
    filter(assessment_date == max(assessment_date)) %>%
    ungroup() %>%
    mutate(is_active = is_present,
           is_active = ifelse(!is.na(inaccessible) & inaccessible == "long term", FALSE, is_active),
           is_active = ifelse(!is.na(inaccessible) & inaccessible == "short term", TRUE, is_active),
           is_active = ifelse(!is.na(not_measurable) & not_measurable == "long term", FALSE, is_active),
           is_active = ifelse(!is.na(not_measurable) & not_measurable == "short term", TRUE, is_active)) %>%
    select(point_code, grts_ranking, grts_ranking_draw, sac_streams, legacy_site, type, segment_id, is_active, date = assessment_date) %>%
    mutate(sac_streams = factor(ifelse(segment_id == "1_4_772", "outside", as.character(sac_streams))),
           legacy_site = FALSE,
           type = "3260",
           type = factor(type, levels = levels(types$type)))

```

```{r}
mhq_streams_refpoints_update %>%
    write_vc(file = "mhq_streams_refpoints",
             digits = 20,
             root = "../../data/mhq_streams",
             sorting = c("grts_ranking", "point_code"))
```


## Update mhq_streams_measurements

```{r}

data_path <- fileman_up("n2khab-mhq-data")

header_mhq_aq <- read_vc(root = file.path(data_path, "processed/inboveg_mhq_aq"), file = "header_mhq_aq")

header_mhq_aq <- header_mhq_aq %>%
    select(measurement_date = vague_date_begin, field_code = location, recording_givid, user_reference) %>%
    mutate(measurement_date = as.Date(measurement_date))

location_code <- admin_streams %>%
    filter(measured) %>%
    mutate(field_code = str_trim(field_code)) %>%
    mutate(segment_id = ifelse(field_code == "Voer_08", "1_4_31", 
                                ifelse(field_code == "BolBe_08", "1_4_491",
                                       ifelse(field_code == "SchNe_04", "1_4_565",
                                               ifelse(field_code == "GroNe_12", "1_4_1602", 
                                                      ifelse(field_code == "Stiemer_02", "1_4_1497", 
                                                             ifelse(field_code == "KleBe_01", "1_7_171136",
                                                                    ifelse(field_code == "KaaBe_01a", "1_4_772", segment_id)))))))) %>%
    
    distinct(field_code, segment_id)

measurements <- mq_streams_assessments_update %>%
    filter(measured) %>%
    left_join(location_code, by = "segment_id") %>%
    mutate(fieldwork_team = "inbo") %>%
    select(fieldwork_team, point_code, segment_id, field_code, measurement_date = assessment_date) %>%
    left_join(header_mhq_aq, by = c("measurement_date", "field_code"))

check <- admin_streams_orig %>%
    filter(klasse == "mon3260") %>%
    distinct(Veldcode, sampling_unit_code, klasse, `datum veldbezoek`) %>%
    mutate(check = Veldcode %in% measurements$field_code)

overview <- measurements %>%
    left_join(select(mhq_streams_popunits, point_code, sac_streams), by = "point_code") %>%
    mutate(cycle = ifelse(year(measurement_date) < 2022, 1, 2)) %>%
    group_by(cycle, sac_streams) %>%
    summarise(n_measured = n_distinct(point_code)) %>%
    ungroup()

sac_streams_seg <- mhq_streams_popunits %>%
    distinct(segment_id, sac_streams)

check2 <- admin_streams %>%
    filter(year(assessment_date) < 2022) %>%
    left_join(sac_streams_seg, by = "segment_id") %>%
    mutate(in_1_7 = segment_id %in% samplingframe_habitatstreams_1_7$segment_id)

measurements %>%
    rename(db_ref = field_code) %>%
    write_vc(file = "mhq_streams_measurements",
             root = "../../data/mhq_streams",
             sorting = c("measurement_date", "point_code"))

```


