
# Progress evaluation for inbo fieldwork team

## Number of valid measured sampling units per type

```{r}
refpoints_validity <- mhq_terr_refpoints_validity %>%
    select(point_code, sampling_unit_code, assessment_date = date, is_valid, is_valid_type, is_valid_refpoint)

measurements <- mhq_terr_refpoints_update %>%
    left_join(mhq_terr_assessments_update, by = c("point_code", "type_target"), suffix = c("_orig", "_measured")) %>%
    filter(is_type_target | (is.na(is_type_target) & assessment_source == "field assessment")) %>%
    left_join(mhq_terr_measurements, c("point_code", "assessment_date", "type_observed", "sampling_unit_code")) %>%
    left_join(refpoints_validity, by = c("point_code", "sampling_unit_code", "assessment_date")) %>%
    filter(is.na(is_valid) | is_valid)

# measurements <- mhq_terr_refpoints_check_update %>%
#     full_join(mhq_terr_assessments_update, by = c("point_code", "type_target"), suffix = c("_orig", "_measured")) %>%
#     filter(is_type_target | (is.na(is_type_target) & assessment_source == "field assessment")) %>%
#     #left_join(mhq_terr_measurements, c("point_code", "assessment_date", "type_observed", "sampling_unit_code")) %>%
#     left_join(refpoints_validity, by = c("point_code", "sampling_unit_code", "assessment_date")) %>%
#     filter(is.na(is_valid) | is_valid)

assessments <- mhq_terr_refpoints_update %>%
    left_join(mhq_terr_assessments_update, by = c("point_code", "type_target"), suffix = c("_orig", "_measured")) %>%
    filter(!is.na(assessment_date))


```


For now we condsider all sampling units for which the validity is currently unknown as valid.

## Check 6410_ve

Do to changes in habitat definition most of the measurents of 6410_ve should actually be 6410_mo. This can be seen in the table below in which 'bwk_description' provides the habitat description according to the 2020 version of the habitatmap.

```{r}
habitatmap <- read_habitatmap_terr()
habitatmap_pol <- habitatmap$habitatmap_terr_polygons

measurements_6410_ve <- measurements %>%
  filter(type_observed == "6410_ve") %>%
  st_as_sf(coords = c("x_measured", "y_measured"), crs = 31370) %>%
  st_join(habitatmap_pol) %>%
  st_drop_geometry() %>%
  select(point_code, type_target, type_observed, assessment_date, type_cover_square, distance, bwk_description = description) %>%
  mutate(changesubtype_to_mo = str_detect(bwk_description, "_mo"))

measurements_6410_ve %>%
  kable() %>%
  kable_styling()
```

We will therefore change 6410_ve to 6410_mo both for type_target and type_observed if the habitatmap indicates that 6410_mo is present in the ref_point. This has still to be modified in INBOVEG.

```{r}
measurements_6410_ve_modify <- measurements_6410_ve %>%
  filter(changesubtype_to_mo)

measurements <- measurements %>%
  mutate(type_target = ifelse(type_target == "6410_ve" & point_code %in% measurements_6410_ve_modify$point_code, "6410_mo", as.character(type_target)),
         type_observed = ifelse(type_observed == "6410_ve" & point_code %in% measurements_6410_ve_modify$point_code, "6410_mo", as.character(type_observed)))
```

```{r}
measurements_overview_type <- measurements %>%
    group_by(fieldwork_team, type_target, sac) %>%
    summarise(n_measured = n_distinct(sampling_unit_code),
              n_measured_check1 = n_distinct(point_code),
              n_measured_check2 = n()) %>%
    ungroup()

# all.equal(measurements_overview_type$n_measured, 
#           measurements_overview_type$n_measured_check1,
#           measurements_overview_type$n_measured_check2)

check <- measurements_overview_type %>%
    filter((n_measured != n_measured_check1) | n_measured != n_measured_check2)

measurements_overview_type <- measurements_overview_type %>%
  select(fieldwork_team, type = type_target, sac, n_measured)

measurements_overview_year <- measurements %>%
    mutate(year = year(assessment_date)) %>%
    group_by(fieldwork_team, year) %>%
    summarise(n_measured = n_distinct(sampling_unit_code),
              n_measured_check1 = n_distinct(point_code),
              n_measured_check2 = n()) %>%
    ungroup()

assessments_table_new %>%
    #filter(lsvi_measurement | is.na(lsvi_measurement)) %>%
    filter(is_type_target | is.na(is_type_target)) %>%
    nrow()

# new_sampling_units %>%
#   anti_join(measurements, by = c("point_code", "type_target"))
```



## Sample size

We will use adatpted sample size as decided in 2021:

+ sample size per type for the whole of Flanders: 80
+ oversampling within sac: add extra sampling units until a number of 80 is reached
+ oversampling within subtypes: add extra sampling units to reach minimum 40 per subtype

```{r}
fileman_up("n2khab-mhq-design")

samplesize_mhq_detailed <- read_vc(root = fileman_up("n2khab-mhq-design"), "design-strategy/output/samplesize_mhq_terr_v2021")

samplesize_mhq <- samplesize_mhq_detailed %>%
      mutate(samplesize = ceiling(n_finite_flanders + n_extra_sac + n_extra_subtype),
             samplesize_type = ceiling(n_finite_flanders + n_extra_sac),
             extra_subtype = ceiling(n_extra_subtype),
         n_visits = ceiling((n_finite_flanders + n_extra_sac + n_extra_subtype)/prop_subtype_sac),
         detection_rate_expected = prop_subtype_sac) %>%
  select(type,area_subtype_sac, sac, samplesize, samplesize_type, extra_subtype,  n_visits, detection_rate_expected) %>%
  mutate(type = ifelse(type == "2190_overig", "2190", as.character(type)))

progress_mhq <- samplesize_mhq %>%
  filter(samplesize > 0) %>%
  left_join(measurements_overview_type, by = c("type", "sac")) %>%
  mutate(fieldwork_team = ifelse(type %in% c("2190_mp", "2190", "6230_hnk", "7140_oli", "6410_ve"), "inbo", fieldwork_team),
         n_measured = ifelse(is.na(n_measured), 0, n_measured),
         progress = round(n_measured/samplesize * 100, 0),
         n_todo = pmax(0, samplesize - n_measured),
         n_visits_todo = ceiling(n_todo/detection_rate_expected)) %>%
  mutate(type_group = ifelse(str_sub(type, 1, 2) == "21", "dunes", "grasslands"))

progress_mhq_inbo <- progress_mhq %>%
  filter(fieldwork_team == "inbo")  %>%
  select(type_target = type, sac, n_todo, n_visits_todo, detection_rate_expected) 

overview_type_group <- progress_mhq %>%
  filter(fieldwork_team == "inbo") %>%
  group_by(type_group) %>%
  summarise(n_measured = sum(n_measured),
            n_todo = sum(n_todo)) %>%
  ungroup()

```

## Progress

The table below gives an overview of the progress per type.

```{r}
progress_mhq%>%
  filter(fieldwork_team == "inbo")  %>%
  select(type, sac, samplesize, n_measured, "progress (%)" = progress, n_todo) %>%
  kbl() %>%
  kable_styling()
```

Following table gives a summary per type category.

```{r}
overview_type_group %>%
  kbl() %>%
  kable_styling(full_width = FALSE)
```



# Sample selection for inbo fieldwork team

```{r, eval=FALSE}

mhq_refpoints_todo_potential <- read_vc(root = fileman_up("n2khab-mhq-design"), file = "design-strategy/output/mhq_refpoints_todo_potential")

mhq_refpoints_todo_potential_update <- mhq_refpoints_todo_potential %>%
    anti_join(assessments_table_new, by = c("point_code")) 

check <- mhq_refpoints_todo_potential %>%
    inner_join(assessments_table_new, by = c("point_code")) %>%
    filter(status == "new")

mhq_selection_inbo <- mhq_refpoints_todo_potential_update %>%
  inner_join(progress_mhq_inbo, by = c("type_target", "sac")) %>%
  group_by(type_target, sac) %>%
  mutate(ranking_abs = rank(grts_ranking_draw)) %>%
  ungroup() %>%
  mutate(select = ranking_abs <= n_visits_todo,
         reserve = ranking_abs > n_visits_todo & ranking_abs < (n_visits_todo * 10))

mhq_selection_inbo_select <- mhq_selection_inbo %>%
  filter(select)

check <- mhq_selection_inbo_select %>%
  group_by(sac, type_target) %>%
  summarise(n_selected = n_distinct(sampling_unit_code),
            n = n()) %>%
  ungroup() %>%
  left_join(progress_mhq_inbo, by = c("sac", "type_target"))
```

+ We will update the selection of sampling units based on the latest version of the habitat map.

+ we will keep the previously selected legacy sampling units (PINK) for the dune habitats that have not been assessed yet

## Legacy sampling units for dune habitats (PINK)

```{r}
refpoints_legacy_todo <- mhq_terr_refpoints_update %>%
    left_join(mhq_terr_assessments_update, by = c("point_code", "type_target"), suffix = c("_orig", "_measured")) %>%
    filter(is.na(assessment_date)) %>%
    filter(legacy_site) %>%
  filter(str_sub(type_target, 1, 2) == "21")

assessments_pink_done <- assessments %>%
  filter(legacy_site) %>%
  filter(str_sub(type_target, 1, 2) == "21")

pink_todo <- refpoints_legacy_todo %>%
  group_by(type_target, sac) %>%
  summarise(n_pink = n_distinct(sampling_unit_code)) %>%
  ungroup()

todo_mhq_inbo <- progress_mhq_inbo %>%
  select(type_target, sac, n_todo) %>%
  left_join(pink_todo, by = c("type_target", "sac")) %>%
  mutate(n_pink = ifelse(is.na(n_pink), 0, n_pink),
        n_todo_new = pmax(0, n_todo - n_pink),
        n_todo_pink = pmin(n_todo, n_pink))
```

The table below shows for each dune habitat type:

+ n_todo: the number of sampling units still needed to reach the desired sample size
+ n_pink: the number of available pink sampling units that have not yet been assessed
+ n_todo_new: the number of new sampling units that we will have to select
+ n_todo_pink: the number of available pink sampling units that will be assessed (only for 2130_hd this is smaller than the available sampling units)


```{r}
todo_mhq_inbo %>%
  filter(str_sub(type_target, 1, 2) == "21") %>%
  kable() %>%
  kable_styling()
```

```{r}
refpoints_legacy_todo_selected <- refpoints_legacy_todo %>%
  left_join(todo_mhq_inbo, by = c("type_target", "sac")) %>%
  group_by(type_target, sac) %>%
  mutate(ranking_rel = rank(grts_ranking_draw),
         selection = ranking_rel <= n_todo_pink * 1.2,
         x_coord = x_orig,
         y_coord = y_orig) %>%
  filter(selection) %>%
  mutate(replacement_rule = "existing PINK PQ: do not replace",
         phab = 100,
         n_type_target = 1,
         type_target_all = type_target) %>%
  st_as_sf(coords = c("x_orig", "y_orig"), crs =31370) %>%
  st_join(habitatmap_pol) %>%
  st_drop_geometry() %>%
  select(sampling_unit_code, point_code, sac, type_target, phab, n_type_target, type_target_all, grts_ranking, grts_ranking_draw, ranking_rel, x = x_coord, y = y_coord, polygon_id, description_orig, replacement_rule)
  
```


## Selection of new sampling units using an equal probability sample

+ To every sampling unit in the sample frame we assign a grts_ranking and a random number between 0 and 100.

+ The sample is drawn in two steps

  + First all sampling units are selected for which the random number is lower than the phab value of the polygon in which the sampling unit is situated.

  + Next we select the desired number of sampling units according to the grts_ranking.

+ If a selected sampling unit does not contain the target type it has to be replaced with a sampling unit within the same polygon. 

```{r}
grts_habmap <- read_vc(file = "design-strategy/output/samplingframe_habitatterr_points", root = fileman_up("n2khab-mhq-design"))

raster_runif <- raster::raster(file.path(fileman_up("n2khab_data"), "20_processed/raster_runif/raster_runif.tif"))
  
grts_habmap_types <- grts_habmap %>%
  left_join(habitatmap$habitatmap_terr_types, by = "polygon_id") %>%
  filter(!is.na(type)) %>%
  filter(phab > 10) %>%
  group_by(type, sac) %>%
  mutate(ranking_rel = rank(grts_ranking)) %>%
  ungroup()

grts_habmap_types_top1000 <- grts_habmap_types %>%
  group_by(type, sac) %>%
  top_n(1000, desc(grts_ranking)) %>%
  ungroup() 

habitatmap_grts_ranking <- grts_habmap_types_top1000 %>%
  distinct(grts_ranking, x, y) %>%
  arrange(grts_ranking) %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  mutate(select_prob = raster_runif[as(., "Spatial")]) %>%
  st_drop_geometry()

grts_habmap_types_top1000 <- grts_habmap_types_top1000 %>%
  left_join(habitatmap_grts_ranking, by = "grts_ranking")

mhq_refpoints_todo_potential <- grts_habmap_types_top1000 %>%
  anti_join(assessments, by = "grts_ranking") %>%
  filter(! grts_ranking %in% refpoints_legacy_todo_selected$grts_ranking_draw) %>%
  mutate(point_code = str_c(grts_ranking, "_1"),
         grts_ranking_draw = grts_ranking,
         sampling_unit_code = str_c(grts_ranking, "_", type),
         type_target = type)

mhq_refpoints_new_inbo <- mhq_refpoints_todo_potential %>%
  mutate(type_target = ifelse(type_target == "2190_overig", "2190", as.character(type_target))) %>%
  inner_join(todo_mhq_inbo, by = c("type_target", "sac")) %>%
  mutate(select_refpoint = select_prob < phab/100) %>%
  filter(select_refpoint) %>%
  group_by(type_target, sac) %>%
  mutate(ranking_rel = rank(grts_ranking_draw),
         max_grts_ranking = max(grts_ranking_draw),
         x_coord = x,
         y_coord = y) %>%
  ungroup() %>%
  mutate(select = ranking_rel <= n_todo_new * 1.2) %>%
  filter(select) %>%
  group_by(point_code) %>%
  mutate(n_type_target = n(),
         type_target_all = str_c(type_target, collapse = " + ")) %>%
  ungroup() %>%
  left_join(st_drop_geometry(habitatmap_pol), by = "polygon_id") %>%
  # st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  # st_join(habitatmap_pol) %>%
  # st_drop_geometry() %>%
  mutate(replacement_rule = "new sampling unit: replace within polygon if type target is not present") %>%
  select(sampling_unit_code, point_code, sac, type_target, phab, n_type_target, type_target_all, grts_ranking_draw, max_grts_ranking, ranking_rel, x = x_coord, y = y_coord, polygon_id, description_orig, replacement_rule) %>%
  arrange(type_target, sac)
```

## Selection of potential replacement sampling units

For all habitatmap polygons that contain newly selected sampling units, we select potential replacement sampling units. 
We do this by selecting all grts_master cells within the polygon excluding the grts_master cells that have already been assessed.

Note that no replacement is performed for the PINK PQ's.

```{r}

polygons_type_target <- mhq_refpoints_new_inbo %>%
  group_by(polygon_id) %>%
  summarise(type_target_all = str_c(unique(type_target), collapse = " + ")) %>%
  ungroup()

mhq_refpoints_new_inbo_polygons <- habitatmap_pol %>%
  filter(polygon_id %in% mhq_refpoints_new_inbo$polygon_id) %>%
  left_join(polygons_type_target, by = "polygon_id") %>%
  select(polygon_id, description_orig, type_target_all)

mhq_refpoints_new_inbo_polygons %>%
  rename(pol_id = polygon_id, descr = description_orig, type_all = type_target_all) %>%
  st_write("../../output/polygons_todo_inbo_versie2021-05-04.shp", driver = "ESRI Shapefile",
         delete_layer = TRUE)

# grts_ranking_assessed <- mhq_terr_assessments_update %>%
#   filter(!is.na(assessment_source)) %>%
#   left_join(select(mhq_terr_refpoints_update, point_code, grts_ranking, grts_ranking_draw), by = c("point_code")) %>%
#   group_by(point_code, grts_ranking, grts_ranking_draw, assessment_source, x, y) %>%
#   summarise(type_target_all = str_c(type_target, collapse = " + ")) %>%
#   ungroup() %>%
#   mutate(grts_ranking = ifelse(is.na(grts_ranking), grts_ranking_draw, grts_ranking)) %>%
#   st_as_sf(coords = c("x", "y"), crs = 31370) %>%
#   st_join(habitatmap_pol)
  
mhq_replacements <- grts_habmap %>%
  # distinct(grts_ranking, x, y) %>%
  # mutate(x_coord = x,
  #        y_coord = y) %>%
  # st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  # st_join(habitatmap_pol) %>%
  # st_drop_geometry() %>%
  filter(polygon_id %in% mhq_refpoints_new_inbo$polygon_id) %>%
  filter(! grts_ranking %in% assessments$grts_ranking) %>%
  group_by(polygon_id) %>%
  mutate(rank_inpolygon = rank(grts_ranking)) %>%
  ungroup() %>%
  select(polygon_id, grts_ranking, rank_inpolygon, x, y) %>%
  arrange(polygon_id, rank_inpolygon)

write_csv2(mhq_replacements, "../../output/replacements_inbo_versie2021-05-04.csv")

max_grts_ranking <- mhq_refpoints_new_inbo %>%
  group_by(type_target, sac) %>%
  summarise(max_grts_ranking = max(grts_ranking_draw)) %>%
  ungroup()

write_csv2(max_grts_ranking, "../../output/maxranking_inbo_versie2021-05-04.csv")
  
```

## Combine

We combine the selected PINK legacy sampling units and the newly selected sampling units.

```{r}
mhq_refpoints_inbo_todo <- mhq_refpoints_new_inbo %>%
  bind_rows(refpoints_legacy_todo_selected) %>%
  arrange(type_target, sac, grts_ranking_draw) %>%
  select(-grts_ranking)

mhq_refpoints_inbo_todo %>%
  write_csv2("../../output/samplingunits_todo_inbo_versie2021-05-04.csv")
```

```{r}
check <- mhq_refpoints_inbo_todo %>%
  group_by(type_target, sac) %>%
  summarise(n = n(),
            n_check = n_distinct(sampling_unit_code)) %>%
  ungroup() %>%
  left_join(todo_mhq_inbo, by = c("type_target", "sac"))
```



## Check the result

+ The figures below show the overall distribution of the phab values of the selected sampling units and the distribution per type     
```{r}
mhq_refpoints_inbo_todo %>%
  ggplot(aes(x = phab)) +
  geom_histogram(bins = 10)
```


```{r}
mhq_refpoints_inbo_todo %>%
  ggplot(aes(x = phab)) +
  geom_histogram(bin = 10) + 
  facet_wrap(~type_target, scales = "free_y")
```

+ The figure below shows the distribution of the number of replacement points per polygon

```{r}
mhq_replacements %>%
  group_by(polygon_id) %>%
  summarise(n_replacement_points = n()) %>%
  ungroup() %>%
  ggplot(aes(x = n_replacement_points)) +
  geom_histogram()
```


### Example for type 2160

```{r}
type_example <- "2160"

sample_example <- mhq_refpoints_inbo_todo %>%
  filter(type_target == type_example) %>%
  group_by(polygon_id) %>%
  mutate(points_per_polygon = n()) %>%
  ungroup() %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  st_transform(crs = 4326) 

replacements_example <- mhq_replacements %>%
  filter(polygon_id %in% sample_example$polygon_id) %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  st_transform(crs = 4326) 

# assessed_example <- assessments %>%
#   filter(polygon_id %in% sample_example$polygon_id) %>%
#   st_transform(crs = 4326) 

polygons_example <- mhq_refpoints_new_inbo_polygons %>%
  filter(polygon_id %in% sample_example$polygon_id) %>%
  mutate(show_label = str_c(polygon_id, ": ", description_orig)) %>%
  st_transform(crs = 4326) 
  
```



```{r, eval = FALSE}
leaflet(sample_example) %>%
  addTiles() %>%
  #addProviderTiles(providers$GeoportailFrance.orthos) %>%
  addPolygons(data = polygons_example, label = ~show_label) %>%
  addMarkers() %>%
  addCircleMarkers(data = replacements_example, color = "black", radius = 1, label = ~rank_inpolygon) %>%
  addCircleMarkers(data = assessed_example, label = ~type_target_all, color = "yellow")


```

### Small polygon: easy

+ the marker shows a selected sampling unit for type 2160 in a habitatmap polygon with 40% 2160

+ the yellow circles are all sampling units that have not yet assessed, within the polygon 

+ if type 2160 is not present at the selected sampling unit the other sampling units are checked in the correct order


```{r }
polygon <- "595937_v2014"

polygon_example <- polygons_example %>%
  filter(polygon_id == polygon)

pol_sample_example <- sample_example %>%
  filter(polygon_id == polygon)

pol_replacements_example <- replacements_example %>%
  filter(polygon_id == polygon)


leaflet(pol_sample_example, width = 800, height = 600) %>%
  #addTiles() %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addPolygons(data = polygon_example, label = ~show_label) %>%
  addMarkers() %>%
  addCircleMarkers(data = pol_replacements_example, color = "yellow", radius = 3, label = ~rank_inpolygon,
                   labelOptions = labelOptions(noHide = TRUE))

```

### Larger polygon

A larger polygon with 12% 2160.

+ The red circles have already been assessed (for other types).

+ To find a suitable replacement, we will first have to map 2160 in the polygon, and then select the sampling unit within 2160 with the lowest rank

+ Notice that the first replacement point has a lower than the first selected sampling unit. This is caused by the sampling procedure used.


```{r}
polygon <- "645479_v2020"

polygon_example <- polygons_example %>%
  filter(polygon_id == polygon)

pol_sample_example <- sample_example %>%
  filter(polygon_id == polygon)

pol_replacements_example <- replacements_example %>%
  filter(polygon_id == polygon)

# pol_assessed_example <- assessed_example %>%
#   filter(polygon_id == polygon)

sd <- SharedData$new(pol_replacements_example)

filter_slider("rank_inpolygon", "Relative ranking in polygon", sd, column = ~rank_inpolygon, step = 1, width = 250)

leaflet(pol_sample_example, width = 800, height = 600) %>%
  #addTiles() %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%

  addPolygons(data = polygon_example, label = ~show_label) %>%
  addCircleMarkers(data = sd, color = "yellow", radius = 3, label = ~rank_inpolygon,
                   labelOptions = labelOptions(noHide = TRUE)) %>%
  # addCircleMarkers(data = pol_assessed_example, color = "red", radius = 3) %>%
  addMarkers(popup = ~type)
  
```

